---
title: "final_data_summary"
author: "Jessica Chung /Melissa Carew"
date: "01/03/2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This final section enables data to be arranged into a final table for future analysis.

# producing final data summary

Starting from here will need the thershold list and the sample_df from Cache03
```{r}
load(here::here("plants/cache/03.RData"))
#save(thresholds_list, file = here("cache/thresholds_list.RData"))
#saveRDS(samples_df, file = here("plants/samples_df.rds"))

# clear Global environment
#load(here::here("plants/results/classification/thresholds/thresholds_list.rds"))
#load(here::here("cache/samples_df.rds"))
```

# load results files from previuos steps
```{r}
zeale_lulu <- readRDS(here("bat_diet/results/lulu/zeale_lulu.rds"))
#zeale_classifications <- read_qza(here("plants/results/classification/zeale_classification.qza"))
zeale_seq <- read_qza(here("bat_diet/results/dada2/zeale_representative_sequences.qza"))
#thresholds_list <- readRDS(here("plants/cache/thresholds_list.rds"))
#samples_df <- readRDS(here::here("plants/cache/samples_df.rds"))
```

# check classification data for zeale amplicon
```{r}
zeale_classifications$data %>% head(100)
```

```{r}
# Pull out table of read counts with LULU filter applied
zeale_lulu$curated_table
```

```{r}
# merge zeale_classifications with thresholds
zeale_df <- merge(zeale_classifications$data, thresholds_list[["zeale"]], by.x="Feature.ID", by.y="feature-id")
```

```{r}
zeale_seq$data %>% head
```


```{r}
trnl_seq$data %>% head
```

```{r}
zeale_df %>% arrange(desc(max_p_identity)) %>% head(100)
```

```{r}
table(zeale_df$Taxon == "Unassigned")
```

```{r}
zeale_df %>% filter(Taxon == "Unassigned") %>% pull(max_p_identity) %>% table
```




```{r}
# for big dataset
zeale_search_results <- read_tsv(here("plants/results/classification/preliminary/zeale_vsearch.tsv"))````

```{r}
zeale_search_results$data %>% dim
```

## Combine workflow outputs

# fix sample names

This can be changed with future feedback.  

We could use https://docs.qiime2.org/2023.5/plugins/available/feature-table/filter-features-conditionally/
if we want simple filtering, but if we want to do something complex, better to do it manually.


```{r}
all_tables <- list(
  zeale=zeale_lulu$curated_table)
```


```{r}
# Get sample names and groups
sample_groups <- samples_df$factor
names(sample_groups) <- samples_df$sample_name
sample_groups
```

# combining read tables for the zeale rbcl and trnl amplicons
```{r}
# make them dataframes
zeale_read_tab <- all_tables$zeale

# Apply the function to the column names
names(zeale_read_tab) <- remove_prefixes(names(zeale_read_tab))
```


```{r}
#merge zeale and trnl reads tables into one dataframe
all_read_df <-  rbind(zeale_read_tab)

# covert row headers to column and then name column 1
all_read_df <- rownames_to_column(all_read_df, var = "asv_code")
```

#Combining read classification tables for the zeale and trnl amplicons
```{r}
# # bind zeale rbcl and trnl classification data
# all_classif_df <-  rbind(zeale_df, rbcl_df, trnl_df)
# 
# # rename column 1
# colnames(all_classif_df)[1] <- "asv_code"
# 
# # join into a singe dataframe
# all_readclass_df <- left_join(all_classif_df, all_read_df, by = "asv_code")
```

# add in ASV sequence data
```{r}
# put sequences in a dataframe
zeale_asvseq <- data.frame(zeale_seq$data)

# convert row headers to column and then name column 1
zeale_asvseq <- rownames_to_column(zeale_asvseq, var = "asv_code")

# add amplicon information
zeale_asvseq  <- mutate(zeale_asvseq, amplicon = "zeale")

# align column names for merging
zeale_asvseq <- zeale_asvseq %>% dplyr::rename(asv_seq = zeale_seq.data)

# merge data into a single dataframe
all_asvseq <-  rbind(zeale_asvseq)
```

```{r}
# merge ALL data into a single dataframe
final_df <- left_join(all_read_df, all_asvseq, by = "asv_code")

# remove columns without data
final_df <- na.omit(final_df)

final_df$Taxon <- gsub("k__", "", final_df$Taxon)
final_df$Taxon <- gsub("p__", "", final_df$Taxon)
final_df$Taxon <- gsub("c__", "", final_df$Taxon)
final_df$Taxon <- gsub("o__", "", final_df$Taxon)
final_df$Taxon <- gsub("f__", "", final_df$Taxon)
final_df$Taxon <- gsub("g__", "", final_df$Taxon)
final_df$Taxon <- gsub("s__", "", final_df$Taxon)
final_df$Taxon <- gsub("_", " ", final_df$Taxon)

# place taxonomic data in separate columns
final_df <- separate(final_df, Taxon, into = c("kingdom", "phylum", "class", "order", "family", "genus","species"), sep = ";",  fill = "right")

head(final_df)
```

```{r}
# save dataframe
write.csv(final_df, file = here("bat_diet/zeale_final_data_summary.csv"), row.names = FALSE)
```

```{r}
wd1 <- "~/git/metabarcoding_workflow_MiSeq22/bat_diet"

# read in .csv files with bold barcodeID engine results.
bold_results <- read.csv(paste0(wd1, "/Zeale_asv_bold_search.csv"))

# rename all results to match other dataframes
bold_results <- bold_results %>%
  dplyr::rename(asv_code = Query.ID) %>%
  dplyr::rename(perc_identity = ID.)

#  extract 'bin_url' and place in a new column
bold_results <- bold_results %>%
  mutate(bin_uri = str_extract(PID..BIN., "(?<=:)(.*?)(?=])"))

bold_results <- bold_results %>%
  filter(bin_uri != "ADC3402")
```

```{r}
## summarize the bold data keeping the highest value for ID (note this doesn't always give the lowest taxonomy, just the best matching record). Remove PID..BIN. column so duplicate asv_code can be removed.
bold_results <- bold_results %>%
  select(-"PID..BIN.")

# summarise so there is one  asv_code per row
bold_results_sum <- bold_results %>%
  group_by(asv_code) %>%
  slice_max(order_by = perc_identity, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  distinct(asv_code, .keep_all = TRUE)

final_df <- bold_results_sum %>%
  select(asv_code, Phylum, Class, Order, Family, Genus, Species, bin_uri) %>%
  left_join(final_df, by = "asv_code")

```
```

