---
title: "final_data_summary"
author: "Jessica Chung /Melissa Carew"
date: "01/03/2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This final section enables data to be arranged into a final table for future analysis.

# producing final data summary

# load results files from previuos steps
```{r}
library(dplyr)
zeale_lulu <- readRDS(here("bat_diet/results/lulu/zeale_lulu.rds"))
#zeale_classifications <- read_qza(here("plants/results/classification/zeale_classification.qza"))
zeale_seq <- read_qza(here("bat_diet/results/dada2/zeale_representative_sequences.qza"))
#thresholds_list <- readRDS(here("plants/cache/thresholds_list.rds"))
#samples_df <- readRDS(here::here("plants/cache/samples_df.rds"))
```

# check classification data for zeale amplicon
```{r}
#zeale_classifications$data %>% head(100)
```

```{r}
# Pull out table of read counts with LULU filter applied
zeale_lulu$curated_table
```


```{r}
# merge zeale_classifications with thresholds
#zeale_df <- merge(zeale_classifications$data, thresholds_list[["zeale"]], by.x="Feature.ID", by.y="feature-id")
```

```{r}
zeale_seq$data %>% head
```

```{r}
#zeale_df %>% arrange(desc(max_p_identity)) %>% head(100)
```

```{r}
#table(zeale_df$Taxon == "Unassigned")
```

```{r}
#zeale_df %>% filter(Taxon == "Unassigned") %>% pull(max_p_identity) %>% table
```


```{r}
# for big dataset
#zeale_search_results <- read_tsv(here("plants/results/classification/preliminary/zeale_vsearch.tsv"))
```

```{r}
#zeale_search_results$data %>% dim
```

## Combine workflow outputs

# fix sample names

This can be changed with future feedback.  

We could use https://docs.qiime2.org/2023.5/plugins/available/feature-table/filter-features-conditionally/
if we want simple filtering, but if we want to do something complex, better to do it manually.


```{r}
zeale_table <- list(zeale=zeale_lulu$curated_table)
```


```{r}
# Get sample names and groups
# sample_groups <- samples_df$factor
# names(sample_groups) <- samples_df$sample_name
# sample_groups
```

# combining read tables for the zeale rbcl and trnl amplicons
```{r}
# make them dataframes
zeale_read_tab <- zeale_table$zeale
```

```{r}
# covert row headers to column and then name column 1
zeale_read_tab <- rownames_to_column(zeale_read_tab, var = "asv_code")
```

#Combining read classification tables for the zeale and trnl amplicons
```{r}
# # rename column 1
# colnames(zeale_readclass_df)[1] <- "asv_code"
# 
# # join into a singe dataframe
# zeale_readclass_df <- left_join(zeale_df, zeale_read_tab, by = "asv_code")
```

# add in ASV sequence data
```{r}
# put sequences in a dataframe
zeale_asvseq <- data.frame(zeale_seq$data)

# convert row headers to column and then name column 1
zeale_asvseq <- rownames_to_column(zeale_asvseq, var = "asv_code")

# add amplicon information
zeale_asvseq  <- mutate(zeale_asvseq, amplicon = "zeale")

# align column names for merging
zeale_asvseq <- zeale_asvseq %>% dplyr::rename(asv_seq = zeale_seq.data)

```

```{r}
# merge ALL data into a single dataframe
final_df <- left_join(zeale_read_tab, zeale_asvseq, by = "asv_code")

# remove columns without data
final_df <- na.omit(final_df)

# final_df$Taxon <- gsub("k__", "", final_df$Taxon)
# final_df$Taxon <- gsub("p__", "", final_df$Taxon)
# final_df$Taxon <- gsub("c__", "", final_df$Taxon)
# final_df$Taxon <- gsub("o__", "", final_df$Taxon)
# final_df$Taxon <- gsub("f__", "", final_df$Taxon)
# final_df$Taxon <- gsub("g__", "", final_df$Taxon)
# final_df$Taxon <- gsub("s__", "", final_df$Taxon)
# final_df$Taxon <- gsub("_", " ", final_df$Taxon)
# 
# # place taxonomic data in separate columns
# final_df <- separate(final_df, Taxon, into = c("kingdom", "phylum", "class", "order", "family", "genus","species"), sep = ";",  fill = "right")
# 
# 
print(final_df)
```

```{r}
# save dataframe
write.csv(final_df, file = here("plants/zeale_sample_data_summary.csv"), row.names = FALSE)
```

```{r}

 # read in .csv files with bold barcodeID engine results.
bold_results <- read.csv(here::here("bat_diet/Zeale_asv_bold_search.csv"))

# rename all results to match other dataframes
bold_results <- bold_results %>%
   dplyr::rename(asv_code = Query.ID) %>%
   dplyr::rename(perc_identity = ID.)  %>%
   dplyr::rename(species = Species)  %>%
   dplyr::rename(genus = Genus)  %>%
   dplyr::rename(family = Family)  %>%
   dplyr::rename(order = Order)  %>%
   dplyr::rename(class = Class)  %>%
   dplyr::rename(phylum = Phylum)
  
#  extract 'bin_url' and place in a new column
bold_results <- bold_results %>%
   mutate(bin_uri = str_extract(PID..BIN., "(?<=:)(.*?)(?=])"))

bold_results <- bold_results %>%
  group_by(bin_uri) %>%
  filter(
    # Keep the row if:
    !is.na(species) & species != "" |
    
    # Or if all rows with this bin_uri have no species info
    !any(!is.na(species) & species != "")
  ) %>%
  ungroup()



## summarize the bold data keeping the highest value for ID (note this doesn't always give the lowest taxonomy, just the best matching record). Remove PID..BIN. column so duplicate asv_code can be removed.
bold_results <- bold_results %>%
   select(-"PID..BIN.")

# summarise so there is one  asv_code per row
bold_results_sum <- bold_results %>%
   group_by(asv_code) %>%
   slice_max(order_by = perc_identity, n = 1, with_ties = FALSE) %>%
   ungroup() %>%
   distinct(asv_code, .keep_all = TRUE)

final_df <- bold_results_sum %>%
   select(asv_code, phylum, class, order, family, genus, species, bin_uri, perc_identity) %>%
   left_join(final_df, by = "asv_code")

# save dataframe
#write.csv(final_df, file = here("bat_diet/zeale_final_data_summary.csv"), row.names = FALSE)
```

```{r}
#Add best possible ids
 #read in existing bins
bins <- read.csv(here::here("bat_diet/bins.csv"))
#make a bins list from the bold_results

bins2 <- bold_results %>%
  mutate(
    species = as.character(species),
    bin_uri = as.character(bin_uri)
  ) %>%
  filter(!is.na(species) & species != "") %>%
  group_by(bin_uri) %>%
  slice_max(order_by = perc_identity, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(bin_uri, species)

# pull out record withou a species name
bold_species_na_unique <- bold_results %>%
  filter((is.na(species) | species == ""),        # species missing
         perc_identity >= 97) %>%                 # identity threshold
  select(bin_uri, order, family, genus) %>%     # keep relevant columns
  distinct(bin_uri, .keep_all = TRUE)  

bins_named <- bold_species_na_unique %>%
  mutate(
    bin_uri = trimws(bin_uri),  # Remove any accidental leading/trailing spaces
    species_name = case_when(
      !is.na(genus) & genus != ""   ~ paste0(genus, " sp. B-", bin_uri),
      !is.na(family) & family != "" ~ paste0(family, " sp. B-", bin_uri),
      TRUE                          ~ paste0(order, " sp. B-", bin_uri)
    )
  )
  
# add species names using bins
bins_all <- bind_rows(
  bins,
  bins2 %>%
    filter(!is.na(bin_uri)) %>%
    anti_join(bins, by = "bin_uri"),
 bins_named %>%
    filter(!is.na(bin_uri)) %>%
    select(bin_uri, species = species_name) %>%
    anti_join(bins, by = "bin_uri")
)

final_df <- final_df %>% 
  dplyr::rename(species_old = species) %>%
  left_join(bins_all %>% select(bin_uri, species), by = "bin_uri") %>%
  relocate(species, .after = genus) %>%
  select(-species_old)
```


```{r}
# Update 'species' column to only include names if max threshold is above 97
final_df <-  final_df %>%
  dplyr::mutate(species = if_else(perc_identity < 97, "", species))

# Update 'genus' column to only include names if max threshold is above 95
final_df <-  final_df %>%
  dplyr::mutate(genus = if_else(perc_identity < 95, "", genus))

# Update 'family' column to only include names if max threshold is above 90
final_df <-  final_df %>%
  dplyr::mutate(family = if_else(perc_identity < 90, "", family))

# Update 'order' column to only include names if max threshold is above 80
final_df <-  final_df %>%
  dplyr::mutate(order = if_else(perc_identity < 85, "",order))
```


```{r}
# save dataframe
write.csv(final_df, file = here("bat_diet/bat_diet_final_data_summary.csv"), row.names = FALSE)
```

```{r}
metab_long <- final_df %>%
  pivot_longer(cols = bat1:bat95, names_to = "sample", values_to = "reads") %>%
  filter(reads != 0) #%>%  # Remove rows where value is zero
  #select(-matches("^X(\\.\\d+)?$"))   %>% # Remove unwanted columns named X, X.1, etc.

```

