---
title: "Data_summary_from_bioinformatic_pipeline"
author: "Jessica Chung /Melissa Carew"
date: "01/03/2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This notebook arranges the data from the bioinformatic pipeline into a final table for data filling and filtering (in notebaook '06_filling_taxonomy_PCR_rep_filtering'.

```{r}
library(dplyr)
```

# producing data summary table

Starting from here will need the thersholds list and the sample_df from cache03. If it was not saved individually you will need to load the entire cache
```{r}
#load all of cahche03
#load(here::here("results/cache/03.RData"))

# load thresholds_list form cahche03 
thresholds_list <- readRDS(here::here("results/cache/thresholds_list.rds"))
samples_df <- readRDS(here::here("results/cache/samples_df.rds"))
```

# load results files from previuos steps
These tables are required for assembling final data table
```{r}
short_lulu <- readRDS(here::here("results/lulu/short_lulu.rds"))
short_classifications <- qiime2R::read_qza(here::here("results/classification/short_classification.qza"))
short_seq <- qiime2R::read_qza(here::here("results/dada2/short_representative_sequences.qza"))
```

# check classification data for short amplicon
```{r}
short_classifications$data %>% head(100)
```

```{r}
# Pull out table of read counts with LULU filter applied
short_lulu$curated_table
```

```{r}
# merge short_classifications with thersholds
short_df <- merge(short_classifications$data, thresholds_list[["short"]], by.x="Feature.ID", by.y="feature-id")
```

```{r}
short_seq$data %>% head
```

```{r}
short_df %>% arrange(desc(max_p_identity)) %>% head(100)
```

```{r}
table(short_df$Taxon == "Unassigned")
```

```{r}
short_df %>% filter(Taxon == "Unassigned") %>% pull(max_p_identity) %>% table
```

```{r}
# load classification results for short amplicon
short_search_results <- readr::read_tsv(here::here("results/classification/preliminary/short_vsearch.tsv"))
```

## Combine workflow outputs

# fix sample names (if they have 'X' at the start)
This is the case for sampled from the 10 sites study as the names start with numbers

```{r}
all_tables <- list(
  short=short_lulu$curated_table)
```


```{r}
# Get sample names and groups
sample_groups <- samples_df$factor
names(sample_groups) <- samples_df$sample_name
sample_groups
```

# combining read tables for the short and right amplicons
```{r}
#make them dataframes
short_read_tab <- all_tables$short
```

```{r}
# Do the amplicon datasets contain the same number of samples?
short_read_tab %>% dim
```

```{r}
#merge short and right reads tables into one dataframe
all_read_df <-  short_read_tab

# covert row headers to column and then name column 1
all_read_df <- tibble::rownames_to_column(all_read_df, var = "asv_code")
```

# combining read classification tables for the short and right amplicons
```{r}
# bind short and right classification data
all_classif_df <-  rbind(short_df)

# rename column 1
colnames(all_classif_df)[1] <- "asv_code"

# join into a singe dataframe
all_readclass_df <- left_join(all_classif_df, all_read_df, by = "asv_code")
```

# add in ASV sequence data
```{r}
# put sequences in a dataframe
short_asvseq <- data.frame(short_seq$data)

# covert row headers to column and then name column 1
short_asvseq <- tibble::rownames_to_column(short_asvseq, var = "asv_code")

# add amplicon information
short_asvseq  <- mutate(short_asvseq, amplicon = "short")

# align column names for merging
short_asvseq <- short_asvseq %>% dplyr::rename(asv_seq = short_seq.data)

# merge data into a single dataframe
all_asvseq <-  short_asvseq
```

```{r}
# merge ALL data into a single dataframe
final_df <- left_join(all_readclass_df, all_asvseq, by = "asv_code")

# remove columns without data
final_df <- na.omit(final_df)

final_df$Taxon <- gsub("k__", "", final_df$Taxon)
final_df$Taxon <- gsub("p__", "", final_df$Taxon)
final_df$Taxon <- gsub("c__", "", final_df$Taxon)
final_df$Taxon <- gsub("o__", "", final_df$Taxon)
final_df$Taxon <- gsub("f__", "", final_df$Taxon)
final_df$Taxon <- gsub("g__", "", final_df$Taxon)
final_df$Taxon <- gsub("s__", "", final_df$Taxon)

# place taxonomic data in separate columns
final_df <- tidyr::separate(final_df, Taxon, into = c("kingdom", "phylum", "class", "order", "family", "genus","species"), sep = ";",  fill = "right")

# final tidy up (remove 'NA' and X from start of sample names)
final_df[is.na(final_df)] <- ""
names(final_df) <- sub("X", "", names(final_df))
head(final_df)
```



```{r}
# save dataframe create a directory for results
dir.create(here::here("results/final_data_summaries"))

timestamp <- format(Sys.time(), "%d_%m_%Y")

# Create the file name with the timestamp 
file_name <- paste0(here::here("results/final_data_summaries/MW_short_vsearch_data_summary_"), timestamp, ".csv")

# Save the dataframe with the new file name
write.csv(final_df, file = file_name, row.names = FALSE)
```

