---
title: "Appendix 3. Importing miseq data into Melbourne Water macroinvertebrate database"
format: html
editor_options: 
  chunk_output_type: console
---

## Introduction

This document reads the miSeq data prepared in the first 6 metabarcoding workflow documents. The code in this document consolidates the data, and formats them ready to be imported into the [Melbourne Water macroinvertebrate database](https://tools.thewerg.unimelb.edu.au/mwbugs/) (mwbugs database).  The code in the document begins by making a connection to the main PostgreSQL database on the University of Melbourne Server. For users outside the unimelb network, the database can be accessed using the instructions in Appendix B of Walsh (2025)[^1].

[^1]: Walsh, C.J. (2025) The Melbourne Water macroinvertebrate database: user’s manual. Version 2.1. Melbourne Waterway Research-Practice Partnership Report 20.1b. School of Agriculture, Forest and Ecosystem Sciences, The University of Melbourne, Melbourne. URL: https://tools.thewerg.unimelb.edu.au/mwbugs_manual/.

The first part of the document assembles taxonomic tables from the asv_library compiled in Appendix 2, and generates new taxoncodes from its species.  

```{r}
#| label: setup

source("https://tools.thewerg.unimelb.edu.au/data/mwbugs/bug_database_functions.R")

# Connect to the mwbugs database and load all tables (using the load_all_mwbugs_tables.R 
#  from bug_database_functions.R)

db <- RPostgres::dbConnect(drv = RPostgres::Postgres(), dbname = "mwbugs")

load_all_mwbugs_tables(db)

# Function equivalent to excel's =RIGHT() and =LEFT() functions
right <- function(x, n) {substr(x, nchar(x) - n + 1, nchar(x))}
left <- function(x, n) {substr(x, 1, nchar(x) - n)}

# List of candidate character pairs for filling bugcodes
# avoid confusion of 0s and Os - exclude Os, and exclude Zs which are reserved for unknowns
all_prs <- expand.grid(x = c(0:9,LETTERS[-c(15,26)]), 
                       y = c(0:9,LETTERS[-c(15,26)]), stringsAsFactors = FALSE)
all_prs <- all_prs[order(all_prs$x),]
all_prs$prs <- paste0(all_prs$x, all_prs$y)
# and remove all 0s combined with letters and 00
zeros <- which((grepl("0", all_prs$prs) & 
                  grepl("[A-Z]", all_prs$prs)) | all_prs$pr == "00")
all_prs <- all_prs[-zeros,]
# order with numbers first, and mark numbers
all_prs$number <- 0
# suppressWarnings, because as.numeric warns about introduction of NAs, which this next line deals with
all_prs$number[!is.na(suppressWarnings(as.numeric(all_prs$prs)))] <- 1
all_prs <- all_prs[order(all_prs$number, decreasing = TRUE),]
# order alphabetic codes so that A1 comes before 1A etc...(probably won't use 1A etc)
all_prs <- rbind(all_prs[all_prs$number == 1,],
                 all_prs[grepl("[A-Z]", all_prs$x) & all_prs$number == 0,],
                 all_prs[!grepl("[A-Z]", all_prs$x) & all_prs$number == 0,])

```

1. Import the asv_library as compiled in Appendix 2.

Correct taxonomy names prior to matching to the database

```{r}

# Before we begin, I need to remove 4633 duplicate records already in taxon_spp (an error with the 46-site data?)
taxon_spp <- unique(taxon_spp)

# Bold results table used in Appendix 2
bold_results <- read.csv("~/uomShare/wergStaff/MelCarew/git-data/metabarcoding_workflow/asv_source_files/match_bold_v5.csv")

# Trial run with asv_library taxa (note this creates new bugdata_spp, gen etc from above)
tfill <- as.data.frame(readxl::read_excel("~/uomShare/wergStaff/ChrisW/temp/asv_library_corrected.xlsx",
                       sheet = "tfill"))
tfill <- tfill[tfill$aqu == 1,] #19,706 records

# aqu really only indicates those records that have gone through the full check. Some should still be removed as 
# being not (necessarily) aquatic or macroinvertebrates
tfill <- only_aqu_inverts(tfill, db, TRUE)$include # 19,704...not sure I believe that. Let's come back to it.
# remove dubious_id records
tfill <- tfill[tfill$dubious_id == 0,] #19,675
tfill_spp_bins <- as.data.frame(readxl::read_excel("~/uomShare/wergStaff/ChrisW/temp/asv_library_corrected.xlsx",
                       sheet = "bin_uris"))
tfill_spp_bins <- tfill_spp_bins[tfill_spp_bins$asv_code %in% tfill$asv_code,] #23,213 species/bin combos
bugdata <- unique(tfill[c("kingdom","phylum","class","order","family","genus","species")]) #1,162 unique identifications
bugdata_spp <- bugdata[!is.na(bugdata$species),] #1003
bugdata_gen <- bugdata[is.na(bugdata$species) & !is.na(bugdata$genus),] #96
bugdata_fam <- bugdata[is.na(bugdata$species) & is.na(bugdata$gen) & !is.na(bugdata$family),] #48
bugdata_oth <- bugdata[is.na(bugdata$species) & is.na(bugdata$gen) & is.na(bugdata$family),] #15
bugdata_oth <- bugdata_oth[!is.na(bugdata_oth$order) & 
                             bugdata_oth$order %in% c("Ephemeroptera","Plecoptera",
                                                       "Trichoptera","Anaspidacea"),] #3
#of the last, only solely freshwater orders can be considered definitely aquatic for inclusion in the database

# Some named species are also similarly uncertain:
bugdata_spp_nam <- bugdata_spp[is.na(bugdata_spp$family),]
bugdata_spp_nam1 <- bugdata_spp[is.na(bugdata_spp$genus) & !is.na(bugdata_spp$family),]

# None of the species identified to class or order can be classed as freshwater with certainty 
unique(bugdata_spp_nam$order) #"Enchytraeida","Haplotaxida","Trombidiformes","Diptera","Isopoda","Coleoptera","Plumatellida" 
# and should be removed from bugdata_spp
bugdata_spp <- bugdata_spp[!bugdata_spp$species %in% bugdata_spp_nam$species,] #980

# 6 families in bugdata_spp_nam1 contain only some species with aquatic life-history stages 
bugdata_spp <- bugdata_spp[!bugdata_spp$species %in% 
                              bugdata_spp_nam1$species[bugdata_spp_nam1$family %in% 
                                                 c("Carabidae", "Chrysomelidae","Nanophyidae", 
                                                   "Empididae", "Muscidae","Stratiomyidae")],] #952
# Note: I have included Psychodidae and other families, where a minority of species may not be aquatic
#       and I included Olindidae and Sphaeromatidae - mostly marine families, but with representatives 
#       in lowland river reaches. These records should be checked for feasibility.


# Check all families are in the mwbugs database
all_fams <- unique(c(bugdata_spp$family, bugdata_gen$family, bugdata_fam$family))
all_fams <- all_fams[!is.na(all_fams)] #139
non_match_fams <- all_fams[!all_fams %in% c(taxon_fam$family, 
                            taxon_fam$subfamily[!is.na(taxon_fam$subfamily)])] #0 non-matching
# Check all genera are in the mwbugs database
all_genera <- unique(c(bugdata_spp$genus, bugdata_gen$genus))
all_genera <- all_genera[!is.na(all_genera)]
non_match_gen <-  all_genera[!all_genera %in% taxon_gen$genus] # 0 non-matching
```

The following identifies new species to be added to the database.  As the 46-site data was added prior to finalization of species groups, I will remove the 46-site bugdata altogether, and removing any "B-" species names that ended up being folded into species groups...

```{r}

#Remove all taxa that had been added on the initial problematic delivery of the 46-site data
old_dna_spp <- unique(taxon_spp$species[!is.na(taxon_spp$tempNote) & 
                taxon_spp$tempNote == "Added by CW for Spring 2018 barcoding data"])
old_dna_spp[!old_dna_spp %in% bugdata_spp$species]
# All of these species are likely valid (although no longer in the asv_library) 
# with no reason to remove except for 
to_remove <- c("Chamaedrilus aff. sphagnetorum","'Anatopynia' sp. 3",
               "'Anatopynia' sp. 4")
taxon_spp <- taxon_spp[!taxon_spp$species %in% to_remove,] #4,633, down from 4,635
taxon_all <- taxon_all[!taxon_all$taxon %in% to_remove,] # unchanged

old_dna_morphospp <- unique(morphospp_etc$lowesttaxon[ 
          morphospp_etc$type == "DNA barcoded species without confirmed identification" |
          (!is.na(morphospp_etc$tempNote) & 
             morphospp_etc$tempNote == "Added by CW for Spring 2018 e-DNA data")])
# very few of which remain valid after revision of asv_library - 
# And some of the valid ones belong in taxon_spp. best to remove all and start again
morphospp_etc <- morphospp_etc[!morphospp_etc$lowesttaxon %in% old_dna_morphospp,] #570 down from 954

all_spp <- unique(bugdata_spp$species)  #951
new_spp <- all_spp[!all_spp %in% c(taxon_spp$species, morphospp_etc$lowesttaxon)] #579

# Worth exploring groups as a final check of tfill, I think
groups <- all_spp[grepl(" group", all_spp) | grepl("\\/", all_spp)]
group_members <- data.frame(group = NA, n_bins = NA, bins = NA, species = NA, min_sim = NA)[0,]
for(i in 1:length(groups)){ # takes a minute to run...
  spp_binsi <- tfill_spp_bins[tfill_spp_bins$species == groups[i],]
  binsi <- unique(spp_binsi$bin_uri)
  asvsi <- unique(spp_binsi$asv_code)
  boldi <- bold_results[0,]
  for(j in 1:length(asvsi)){
    boldi <- unique(rbind(boldi, 
                   bold_results[bold_results$asv_code == asvsi[j] & bold_results$bin_uri %in% binsi,]))
  }
  boldi_97 <- boldi[boldi$similarity >= 97,]
  bold_spp <- unique(boldi_97$species) 
  # just those names with >= 2 words
  bold_spp <- bold_spp[sapply(strsplit(bold_spp," "),length) >= 2]
  bold_spp <- bold_spp[!right(bold_spp,4) == " sp."]
  bold_spp <- paste0("('",paste(unique(paste(sapply(strsplit(bold_spp," "),"[[",1), 
                     sapply(strsplit(bold_spp," "),"[[",2))), collapse = "', '"), "')")
  group_members <- rbind(group_members,
                  data.frame(group = groups[i],
                             n_bins = length(binsi),
                             bins = paste0("('",paste( binsi, collapse = "', '"), "')"),
                             species = bold_spp,
                             min_sim = min(boldi$similarity)))
}
# group_members[i,] ; i <- i + 1

# CW manually checked all form_spp names against mwbugs taxon_spp. 
form_spp <- new_spp[!grepl("MC-",new_spp) & !grepl("B-",new_spp) & !grepl(" group", new_spp) &
                    !grepl("/", new_spp)]
form_spp <- form_spp[order(form_spp)]

correct_words <- taxon_spp$species[stringdist::amatch(form_spp, taxon_spp$species, maxDist = Inf)]
data.frame(form_spp, correct_words)

# # Visual inspection revealed equivalences, manually corrected either n the database or in tfill
# "Adversaeschna brevistyla"="Aeshna brevistyla", # the former is current
# "Anax papuensis"="Hemianax papuensis", # the former is current
# "Austrocnephia orientalis"="Paracnephia orientalis") # the former is more current
# "Dicranomyia whitei whitei"="Dicranomyia whitei" # changed above
# "Megogata necopina"="Megogota necopina"  # corrected in database
# "Notalina bifara"="Notalina bifaria" # corrected in next version of App 2 script
# "Offadens nr. confluens"="Offadens confluens" # a diffferent species. Leave it
# "Riekoperla cf. intermedia"="Riekoperla intermedia" # a diffferent species. Leave it
# For the first three, I added entries to the taxon_syns table in the database and changed the species names in taxon_spp and taxon_all in build_mwbugs_db.R

# for identifying families from species names using subfamilies
subfams <- read.csv("~/uomShare/wergStaff/MelCarew/git-data/metabarcoding_workflow/asv_source_files/taxon_subfams_partial.csv")

taxon_spp[!taxon_spp$spcode %in% taxon_all$shortcode,]
taxon_gen[!taxon_gen$spcode %in% taxon_all$shortcode,]
taxon_fam[!taxon_fam$spcode %in% taxon_all$shortcode,]
taxon_rank1[!taxon_rank1$spcode %in% taxon_all$shortcode,]
# All good, no entries in any of these tables not in taxon_all, however, some problems the other way

taxon_all[taxon_all$table == "taxon_spp",][!taxon_all$shortcode[taxon_all$table == "taxon_spp"] %in% taxon_spp$spcode,]
# 4 records, none in the biota table, so delete all
taxon_all <- taxon_all[!taxon_all$shortcode %in% taxon_all$shortcode[taxon_all$table == "taxon_spp"][!taxon_all$shortcode[taxon_all$table == "taxon_spp"] %in% taxon_spp$spcode],]
taxon_all[taxon_all$table == "morphospp_etc",][!taxon_all$shortcode[taxon_all$table == "morphospp_etc"] %in% morphospp_etc$taxoncode,]
x <- biota[biota$taxoncode %in% taxon_all$shortcode[taxon_all$table == "morphospp_etc"][!taxon_all$shortcode[taxon_all$table == "morphospp_etc"] %in% morphospp_etc$taxoncode],]
unique(x$smpcode) # all AL or AM samples, which are going to be removed and replaced, so remove
taxon_all <- taxon_all[!taxon_all$shortcode %in% taxon_all$shortcode[taxon_all$table == "morphospp_etc"][!taxon_all$shortcode[taxon_all$table == "morphospp_etc"] %in% morphospp_etc$taxoncode],]
taxon_all[taxon_all$table == "taxon_gen",][!taxon_all$shortcode[taxon_all$table == "taxon_gen"] %in% taxon_gen$gencode,]
taxon_all[taxon_all$table == "taxon_fam",][!taxon_all$shortcode[taxon_all$table == "taxon_fam"] %in% taxon_fam$famcode,]
taxon_all[taxon_all$table == "taxon_rank1",][!taxon_all$shortcode[taxon_all$table == "taxon_rank1"] %in% taxon_rank1$t1code,]
taxon_spp[!taxon_spp$spcode %in% taxon_all$shortcode,]
taxon_gen[!taxon_gen$spcode %in% taxon_all$shortcode,]
taxon_fam[!taxon_fam$spcode %in% taxon_all$shortcode,]
taxon_rank1[!taxon_rank1$spcode %in% taxon_all$shortcode,]
# OK, all consistent now, so carry on creating new bugcodes...

new_spp <- data.frame(bugcode = NA, species = new_spp, mwbugs_table = NA)
for(i in 1:nrow(new_spp)){
  spi <- new_spp$species[i]
  geni <- strsplit(spi, " ")[[1]][1]
  gencode_i <- taxon_gen$gencode[taxon_gen$genus == geni]
  if(length(gencode_i) > 0){
#    if(substr(gencode_i,1,2) %in% c("IF","IB")) stop("1")
  if(i == 1) new_spp_gencodes <- NULL
  if(i > 1) new_spp_gencodes <- new_spp$bugcode[1:(i-1)][grep(geni, new_spp$species[1:(i-1)])]
  all_geni_codes <- c(taxon_spp$spcode[taxon_spp$gencode == gencode_i],
                      morphospp_etc$taxoncode[substr(morphospp_etc$lowesttaxon,1,6) == geni],
                      new_spp_gencodes)
  used_suffixes <- substr(all_geni_codes,7,8)
  if(geni %in% taxon_gen$genus){
  if(!grepl("MC-",spi) & !grepl("B-",spi) & !grepl(" group", spi) & !grepl("/", spi)){
 # If it is a formal species name, give it the next numeric code.
     used_suffixes <- used_suffixes[used_suffixes %in% all_prs$prs[all_prs$number == 1]]
     new_spp$bugcode[i] <- paste0(gencode_i, all_prs$prs[all_prs$number == 1 & 
                                                     !all_prs$prs %in% used_suffixes][1])
     new_spp$mwbugs_table[i] <- "taxon_spp"
  }else{
  # or else, give it the next alphanumeric code.
     used_suffixes <- used_suffixes[used_suffixes %in% all_prs$prs[all_prs$number == 0]]
     new_spp$bugcode[i] <- paste0(gencode_i, all_prs$prs[all_prs$number == 0 & 
                                                     !all_prs$prs %in% used_suffixes][1])
     new_spp$mwbugs_table[i] <- "taxon_spp"
  }
    }
    }else{
      #If species only identified to family or subfamily...
    if(geni %in% subfams$subfamily & !geni %in% taxon_fam$subfamily){
      geni <- subfams$family[subfams$subfamily == geni & !is.na(subfams$subfamily)]
    }
    if(geni %in% taxon_fam$family){
      famcodei <- taxon_fam$famcode[taxon_fam$family == geni]
      if(geni == "Chironomidae") famcodei <- "QDAZ"
    }else{
    if(geni %in% taxon_fam$subfamily){
      famcodei <- taxon_fam$famcode[!is.na(taxon_fam$subfamily) & taxon_fam$subfamily == geni]
      if(geni == "Chironominae") famcodei <- "QDAJ"
    }
    }
    if(length(famcodei) > 0)
#    if(substr(famcodei,1,2) %in% c("IF","IB")) stop("2")
      if(i == 1) new_spp_fam99codes <- NULL
      if(i > 1) new_spp_fam99codes <- new_spp$bugcode[substr(new_spp$bugcode,1,6) == 
                                             paste0(famcodei,"99")][1:(i-1)]
    allfam99i_codes <- c(taxon_all$shortcode[substr(taxon_all$shortcode,1,6) == 
                                             paste0(famcodei,"99")],
                         new_spp_fam99codes)
    used_suffixes <- substr(allfam99i_codes,7,8)
    used_suffixes <- used_suffixes[used_suffixes %in% all_prs$prs[all_prs$number == 0]]
    new_spp$bugcode[i] <- paste0(famcodei, "99", all_prs$prs[all_prs$number == 0 & 
                                                     !all_prs$prs %in% used_suffixes][1])
    new_spp$mwbugs_table[i] <- "morphospp_etc"
    }
}

# Check this doesn't introduce any duplicate bugcodes
dups <- c(new_spp$bugcode, taxon_all$shortcode)[duplicated(c(new_spp$bugcode, taxon_all$shortcode))]
dups %in% c(taxon_spp$spcode, morphospp_etc$taxoncode)
taxon_spp[taxon_spp$spcode %in% dups[dups %in% c(taxon_spp$spcode, morphospp_etc$taxoncode)],] #No duplicates!

taxon_spp <- rbind(taxon_spp,
                       data.frame(gencode = substr(new_spp$bugcode,1,6),
                            spcode = new_spp$bugcode,
                            species = new_spp$species,
                            tempNote = "added by CW for ARC metabarcoding project",
                            taxonomic_note = NA)[new_spp$mwbugs_table == "taxon_spp",])
taxon_spp <- taxon_spp[order(taxon_spp$spcode),]

taxon_all <- rbind(taxon_all, data.frame(shortcode = new_spp$bugcode,
                                         taxon = new_spp$species, 
                                         table = "taxon_spp"))

morphospp_etc <- rbind(morphospp_etc, 
                       data.frame(taxoncode = new_spp$bugcode,
                                lowesttaxon = new_spp$species,
                                type = "DNA barcoded species",
                                tempNote = "added by CW for ARC metabarcoding project",
                                t1code = substr(new_spp$bugcode,1,2))[new_spp$mwbugs_table == "morphospp_etc",])
taxon_all$table[taxon_all$shortcode %in% 
                  new_spp$bugcode[new_spp$mwbugs_table == "morphospp_etc"]] <-
  "morphospp_etc"

taxon_all <- rbind(taxon_all, data.frame(shortcode = new))

tfill <- dplyr::mutate(tfill, shortcode = NA, .before = max_p_identity)
nocodes <- vector()
for(i in 1:nrow(tfill)){
  if(!is.na(tfill$species[i])){
    if(tfill$species[i] %in% c(taxon_spp$species,morphospp_etc$lowesttaxon)) {
    tfill$shortcode[i] <- ifelse(tfill$species[i] %in% taxon_spp$species,
                    taxon_spp$spcode[taxon_spp$species == tfill$species[i]],
                    morphospp_etc$taxoncode[morphospp_etc$lowesttaxon == tfill$species[i]])
    }
    }else{
   if(!is.na(tfill$genus[i])){
     tfill$shortcode[i] <- taxon_gen$gencode[taxon_gen$genus == tfill$genus[i]]
   }else{
   if(!is.na(tfill$family[i])){
     fami <- tfill$family[i]
     if(grepl("inae",fami)){ 
       tfill$shortcode[i] <- taxon_fam$famcode[taxon_fam$subfamily == tfill$family[i]]
     }else{
     tfill$shortcode[i] <- ifelse(fami == "Chironomidae", "QDAZ",
                                  taxon_fam$famcode[taxon_fam$family == fami])
   }
     }else{
     if(!is.na(tfill$order[i]) & tfill$order[i] == "Ephemeroptera") tfill$shortcode[i] <- "QE"
     if(!is.na(tfill$order[i]) & tfill$order[i] == "Plecoptera") tfill$shortcode[i] <- "QP"
     if(!is.na(tfill$order[i]) & tfill$order[i] == "Trichoptera") tfill$shortcode[i] <- "QT"
     if(!is.na(tfill$order[i]) & tfill$order[i] == "Anaspidacea") tfill$shortcode[i] <- "ON"
     if(!is.na(tfill$order[i]) & tfill$order[i] == "Decapoda") tfill$shortcode[i] <- "OT" 
         # assume Caridea, given earlier ID of these last ones as Atyidae
       # The aquaticness/macroinvertebrateness of all the rest are too uncertain
       # and should be removed from contention for the database
   }
   }
  }
}
# sum(is.na(tfill$shortcode)) # 453 not given a shortcode
unique(tfill$class[is.na(tfill$shortcode)])
unique(tfill$order[is.na(tfill$shortcode)])
# All look worthy of exclusion based on uncertain habit.
tfill <- tfill[!is.na(tfill$shortcode),]  # down to 19,200

taxon_seq <- tfill[c("shortcode","asv_code","asv_sequence","max_p_identity","amplicon")]

asv_bin <- tfill_spp_bins[tfill_spp_bins$asv_code %in% tfill$asv_code,] #22862 out of 23050
asv_bin <- taxon_asv_bin[names(taxon_asv_bin) != "species"]

taxon_seq$taxon <- taxon_all$taxon[match(taxon_seq$shortcode, taxon_all$shortcode)]
asv_bin$taxon <- taxon_seq$taxon[match(asv_bin$asv_code,taxon_seq$asv_code)]
asv_bin$shortcode <- taxon_seq$shortcode[match(asv_bin$asv_code,taxon_seq$asv_code)]
species_bin <- unique(asv_bin[nchar(asv_bin$shortcode) == 8,c("taxon","bin_uri")])
bins_per_spp <- aggregate(species_bin["bin_uri"], 
                          by = list(species = species_bin$taxon), FUN = length)
# revert to original minimalist tables
taxon_seq <- tfill[c("shortcode","asv_code","asv_sequence","max_p_identity","amplicon")]
asv_bin <- asv_bin[c("asv_code","bin_uri","similarity")]

db_revised <- list(taxon_all = taxon_all, taxon_spp = taxon_spp, 
                   morphospp_etc = morphospp_etc, 
                   taxon_seq = taxon_seq,
                   asv_bin = asv_bin)
writexl::write_xlsx(db_revised, "~/uomShare/wergStaff/ChrisW/temp/db_revised.xlsx")
```

The following code pre-dates the finalization of the asv_library, and requires some work....

The following chunk compiles the miseq_samples, samples, and bugdata tables from the source files for 6 miseq runs, checking for and correcting consistency between the tables (with checks back to lab notes for corrections.)

The bugdata is separated into four tables (spp, gen, fam, other) to separate records identified to species (or species group) from those identified to genus, family or a higher level of taxonomy (other).

```{r}
#| echo: true

# Compile all sample data
miseq_dirs <- c(15,16,17,18,19,20)

for(i in 1:length(miseq_dirs)){
  miseq_path <- paste0("~/uomShare/wergStaff/MelCarew/git-data/metabarcoding_workflow/miseq", miseq_dirs[i],"/")
  miseq_sample_info <- paste0("miseq",miseq_dirs[i], "_sample_information.xlsx")
  samples_i <- rio::import_list(paste0(miseq_path,miseq_sample_info))
  miseq_samples_i <- samples_i$miseq_samples
  miseq_samples_i$miseq_run <- miseq_dirs[i]
  #@@ table naming error in miseq15
  names(samples_i)[names(samples_i) == "sample"] <- "samples"
  smpcodes_i <- samples_i$samples
  #@@ column inconsistencies in miseq15, 18, 19, 20
  names(miseq_samples_i)[names(miseq_samples_i) == "comments"] <- "comment"
  names(miseq_samples_i)[names(miseq_samples_i) == "notes"] <- "comment"
  smpcodes_i <- smpcodes_i[,!names(smpcodes_i) %in% c("...7","amplicon")]
  names(miseq_samples_i)[names(miseq_samples_i) == "position"] <- "plate_position"
  names(miseq_samples_i)[names(miseq_samples_i) == "i7 index"] <- "i7_index"
  names(miseq_samples_i)[names(miseq_samples_i) == "i5 index"] <- "i5_index"
  
#  bugdata_file <- dir(paste0(miseq_path, "metabarcoding_workflow/final_modeling_data"))
#  Most recent file - check contents of bugdata_file above to confirm!
  bugdata_file <- paste0("miseq", miseq_dirs[i], 
                         # MC revised miseq18 data after first run of this file identified a problem
                         ifelse(miseq_dirs[i] == 18, 
                                "_final_data_long_format_26_02_2025.csv",
                                "_final_data_long_format_18_02_2025.csv"))
  bugdata_i <- read.csv(paste0(miseq_path, "metabarcoding_workflow/final_modeling_data/", bugdata_file))
  #@@ smpcode inconsistencies in miseq18
  miseq_samples_i$smpcode[miseq_samples_i$smpcode == "388-BF3-373-3"] <- "388-BF3-373"
  #@@ seq_code inconsistencies in miseq15, 18
  miseq_samples_i$seq_code <- gsub("30DNG79574","30DNG318813",miseq_samples_i$seq_code)
  miseq_samples_i$seq_code <- gsub("A22XZZ3644","A22X223644",miseq_samples_i$seq_code)
  
#  bugdata_i$sample <- gsub("A22YZ2161","A22YZ2162", bugdata_i$sample)
  report_i <- data.frame(miseq = miseq_dirs[i], no_bug_rows = nrow(bugdata_i), 
                         no_bug_rows_NA = sum(is.na(bugdata_i$species)), no_new_seqs = NA)
  bugdata_i <- only_aqu_inverts(bugdata_i, db)$include
  
  # Check uniqueness and consistency
  if(sum(duplicated(smpcodes_i$smpcode)) > 0)  stop("1")
  if(sum(duplicated(miseq_samples_i$seq_code)) > 0)  stop("2")
  # Check only entries in smpcodes_i table with NA smpcodes are controls/refs
  # Note tolower() necessary because of inconsistencies with capitalization
  if(sum(!tolower(substr(smpcodes_i$seq_code[is.na(smpcodes_i$smpcode)], 1, 4)) %in% 
          tolower(c("MAR2","Dros","Cont","MART","CO1C","COIC"))) > 0) stop("3")
  smpcodes_i <- smpcodes_i[!is.na(smpcodes_i$smpcode) & 
                             !tolower(substr(smpcodes_i$smpcode,1,4)) %in%
                              tolower(c("MAR2","Dros","cont","MART","CO1C","COIC")),]
  miseq_samples_i <- miseq_samples_i[!tolower(substr(miseq_samples_i$seq_code, 1, 4)) 
                                     %in% tolower(c("MAR2","Dros","Cont","MART","CO1C","COIC")),] 
  #@@ smpcodes missing from reps 4,5,and 6 in miseq15
  missing_smpcodes <- which(is.na(miseq_samples_i$smpcode))
  for(j in missing_smpcodes){
    miseq_samples_i$smpcode[j] <- miseq_samples_i$smpcode[miseq_samples_i$seq_code == 
                                                            paste0(left(miseq_samples_i$seq_code[j],4), "rep1")]
  }
  # #####@@@@ remove miseq18 samples that were re-run in miseq19
    if(miseq_dirs[i] == 18){
  smpcodes_i <- smpcodes_i[smpcodes_i$smpcode != miseq_samples_i$smpcode[miseq_samples_i$seq_code == "A22YZ2162rep3"],]
  miseq_samples_i <- miseq_samples_i[!miseq_samples_i$seq_code %in% c("A22YZ2162rep3","A22YZ2162rep1","A22YZ2162rep2"),]
  bugdata_i <- bugdata_i[!bugdata_i$sample %in% c("A22YZ2162rep3","A22YZ2162rep1","A22YZ2162rep2"),]
}
  # ####@@@@ and non-matching sample codes in the bugdata file
  if(miseq_dirs[i] == 19){
    # Miscoding errors corrected as per emails from Mel 4/12/2024 and 10/12/2024
    miseq_samples_i$smpcode[miseq_samples_i$seq_code == "A22YZ7161rep1"] <- "389-YZ2-162"
    miseq_samples_i$seq_code[miseq_samples_i$seq_code == "A22YZ7161rep1"] <- "A22YZ2162rep1"
    bugdata_i$sample[bugdata_i$sample == "A22BF3373rep1"] <- "A22BF3373rep4"
    bugdata_i$sample[bugdata_i$sample == "A22BF3373rep2"] <- "A22BF3373rep5"
    bugdata_i$sample[bugdata_i$sample == "A22BF3373rep3"] <- "A22BF3373rep6"
  }
  if(miseq_dirs[i] == 20){
    # Miscoding errors corrected as per email from Mel 4/12/2024
    bugdata_i$sample[bugdata_i$sample == "S22ALL58rep2"] <- "S22ALL558rep4"
    bugdata_i$sample[bugdata_i$sample == "S22ALL58rep3"] <- "S22ALL558rep5"
    bugdata_i$sample[bugdata_i$sample == "S22ALL58rep4"] <- "S22ALL558rep6"
  }
  ####@@@@
  if(sum(is.na(match(miseq_samples_i$smpcode,smpcodes_i$smpcode))) > 0) stop("4")
  if(sum(!smpcodes_i$smpcode %in% miseq_samples_i$smpcode) > 0) stop("5")

  # Check and compile bug data
  # Consistency between miseq_samples table and bugdata table.
  # MISSING bugdata: rep2 for this smpcode did not produce enough reads and was 
  # not re-run (see email from Mel 4/12/2024). Delete miseq_sample
  miseq_samples_i <- miseq_samples_i[miseq_samples_i$seq_code != "S21ANS16116rep2",]
  # delete re-run samples from 10-site study (see email from Mel 4/12/2024)
  miseq_samples_i <- miseq_samples_i[!miseq_samples_i$seq_code %in% 
      c("Elut30DNG318813rep1","S21Elut30DNG318813rep2","Elut30DNG318813rep3"),]
  smpcodes_i <- smpcodes_i[smpcodes_i$smpcode != "DNG-7957-4",]
  bugdata_i <- bugdata_i[!bugdata_i$sample %in% 
                           c("Elut30DNG79574rep1", "Elut30DNG79574rep2", "Elut30DNG79574rep3"),]
  ##@@@ Incorrect sitecode in miseq18 and 19 and 20
  miseq_samples_i$seq_code <- gsub("X223644","XZZ3644", miseq_samples_i$seq_code)
  bugdata_i$sample <- gsub("X223644","XZZ3644", bugdata_i$sample)
 # bugdata_i$sample <- gsub("ALL58","ALL558", bugdata_i$sample)
  #@@ spurious Xs at start of sample codes in miseq15
  bugdata_i$sample[substr(bugdata_i$sample,1,1) == "X"] <- 
    substr(bugdata_i$sample[substr(bugdata_i$sample,1,1) == "X"], 2,
           nchar(bugdata_i$sample[substr(bugdata_i$sample,1,1) == "X"]))
  if(sum(!unique(bugdata_i$sample) %in% miseq_samples_i$seq_code) > 0) stop("7")
  # if(sum(!miseq_samples_i$seq_code %in% unique(bugdata_i$sample)) > 0) 
  #   cat(miseq_samples_i$seq_code[!miseq_samples_i$seq_code %in% unique(bugdata_i$sample)],
  #       "miseq",miseq_dirs[i],"\n")
  # The above commented-out bit is ok - it identifies a couple of seq_codes that were re-run
  # and a couple of other problems identified in the next chunk.
  
  # unique(bugdata_i$sample)[!unique(bugdata_i$sample) %in% miseq_samples_i$seq_code]
  # miseq_samples_i$seq_code[!miseq_samples_i$seq_code %in% unique(bugdata_i$sample)]
  # As directed by Mel in email of 4 Dec 2024, remove species with MC_ARCO in their name
  bugdata_i$species[grep("MC-ARCO", bugdata_i$species)] <- ""
  # Remove "environmental sample" records 
  bugdata_i <- bugdata_i[!grepl("environmental", bugdata_i$species) & 
                           !grepl("uncultured zooplankton", bugdata_i$species),]
  # ensure no trailing or leading spaces
  bugdata_i$genus <- stringr::str_trim(bugdata_i$genus)
  # Check consistency with sample tables

  bugdata_i_spp <- bugdata_i[bugdata_i$species != "",]
  # Should all be similarity >= 0.97
  if(sum(bugdata_i_spp$max_p_identity < 0.97) > 0) stop("A")
  bugdata_i_gen <- bugdata_i[bugdata_i$species == "" & bugdata_i$genus != "",]
  bugdata_i_fam <- bugdata_i[bugdata_i$species == "" & bugdata_i$genus == "" & bugdata_i$family != "",]
  bugdata_i_oth <- bugdata_i[bugdata_i$species == "" & bugdata_i$genus == "" & bugdata_i$family == "",]
  # These four tables should contain all records from bugdata_i 
  if(sum(nrow(bugdata_i_spp),nrow(bugdata_i_gen),nrow(bugdata_i_fam),nrow(bugdata_i_oth)) != nrow(bugdata_i)) stop("B")
  if(i == 1){
    bugdata_spp <- bugdata_i_spp
    bugdata_gen <- bugdata_i_gen
    bugdata_fam <- bugdata_i_fam
    bugdata_oth <- bugdata_i_oth
    smpcodes <- smpcodes_i
    miseq_samples <- miseq_samples_i
  }else{
    bugdata_spp <- rbind(bugdata_spp, bugdata_i_spp)
    bugdata_gen <- rbind(bugdata_gen, bugdata_i_gen)
    bugdata_fam <- rbind(bugdata_fam, bugdata_i_fam)
    bugdata_oth <- rbind(bugdata_oth, bugdata_i_oth)
    smpcodes <- rbind(smpcodes, smpcodes_i)
    if(sum(miseq_samples_i$seq_code %in% miseq_samples$seq_code) > 0 )  stop("8")
    miseq_samples <- rbind(miseq_samples, miseq_samples_i)
  }
}
# 2 rogue terrestrials to be deleted.
bugdata_gen <- bugdata_gen[bugdata_gen$genus != "Chrysophtharta",]
bugdata_spp <- bugdata_spp[bugdata_spp$genus != "Eufairmairia",]
```

Some samples were re-run because of insufficient reads (or other reasons), often in a subsequent miseq run.  The following chunk finds such cases (where there are replicate 1, 2, and 3 and 4, 5, and 6), and keeps only replicates 4, 5 and 6.
 
```{r}

# Correct spelling errors (Unnecessary ones kept for tfill run-through, just in case!)
# One lingering spelling error found
# bugdata_spp$species[bugdata_spp$species == "Austrochiltonia sp. MC1"] <- "Austrochiltonia sp. MC-1"
# bugdata_spp$species[bugdata_spp$species == "Tamasia acuta/Tamasia variegata"] <- "Tamasia acuta/variegata"
# bugdata_spp$species <- gsub("ParaLimnophila","Paralimnophila",bugdata_spp$species)
# bugdata_spp$genus[!is.na(bugdata_spp$genus) &  bugdata_spp$genus == "ParaLimnophila"] <- "Paralimnophila"
# bugdata_spp$species <- gsub("Aeschnidae","Aeshnidae",bugdata_spp$species)
# bugdata_spp$species[bugdata_spp$species == "Nososticta solitaris"] <- "Nososticta solitaria"
# # bugdata_spp$family[!is.na(bugdata_spp$family) & bugdata_spp$family == "Trapezidae"] <- "Trapeziidae"
# # # Some Staphylinidae with order = Diptera
# bugdata_spp$order[!is.na(bugdata_spp$family) &  bugdata_spp$family == "Staphylinidae"] <- "Coleoptera"
# # 5 family-level records identified as Unionidae. Genbank match gastropod ~92% - delete
# bugdata_fam <- bugdata_fam[!is.na(bugdata_spp$family) & bugdata_fam$family != "Unionidae",]
# # 6 family-level records identified as Unionidae.  Genbank match gastropod ~92% - delete
# bugdata_fam <- bugdata_fam[!is.na(bugdata_spp$family) &  bugdata_fam$family != "Arcidae",]
# # Unidentified higher-order records of unknown (likely terrestrial) ecology.  Exclude  



sample_reps <- aggregate(miseq_samples$seq_code, 
                         by = list(seq_code = miseq_samples$smpcode), FUN = length)
# sum(!sample_reps$x[i] %in% c(3,6) > 0)  # 0 as expected
  for(j in which(sample_reps$x == 6)){
    miseq_samples <- miseq_samples[!(miseq_samples$smpcode == sample_reps$seq_code[j] &
                                           right(miseq_samples$seq_code, 1) %in% 1:3),]
  }

bugdata_spp <- bugdata_spp[bugdata_spp$sample %in% miseq_samples$seq_code,]
bugdata_gen <- bugdata_gen[bugdata_gen$sample %in% miseq_samples$seq_code,]
bugdata_fam <- bugdata_fam[bugdata_fam$sample %in% miseq_samples$seq_code,]
bugdata_oth <- bugdata_oth[bugdata_oth$sample %in% miseq_samples$seq_code,]

#### Need to remove these seq_codes from the miseq_sample table  
#  (seq_code A22BT2373rep3 & A23YAR36377rep3 not in bugdata_spp table)
#  miseq_samples[!miseq_samples$seq_code %in% unique(bugdata_spp$sample),]
miseq_samples <- miseq_samples[miseq_samples$seq_code %in% unique(bugdata_spp$sample),]

unq_smpcodes <- unique(miseq_samples$smpcode)
# need to associate correct smpcodes with each miseq_samples record. Keep (commonly) incorrect smpcodes as smpcode_mc
miseq_samples <- dplyr::mutate(miseq_samples, smpcode_mc = smpcode, .after = smpcode)
miseq_samples <- dplyr::mutate(miseq_samples, 
                               extraction_name = substr(miseq_samples$seq_code, 1, nchar(miseq_samples$seq_code) - 4), 
                                .after = smpcode_mc)

# Field/planning sample details
a = as.data.frame(readxl::read_excel("~/uomShare/wergData/DNA_Barcoding/DNA barcoding samples processing Spr2021.xlsx"))
samples_a <- a[c("sitecode","sample_code","Actual Date","monthcode","Extraction name","number of riffle")]
samples_a$comment <- NA
samples_a$comment[!is.na(samples_a$`Actual Date_ghd`)] <- "collected by GHD"
samples_a$comment[!is.na(samples_a$`Actual Date_UoM`)] <- "collected by G Hehir, Unimelb"
samples_a$comment[!(is.na(samples_a$`Actual Date_ghd`) & is.na(samples_a$`Actual Date_uom`))] <- "collector unspecified"
names(samples_a) <- c("sitecode","smpcode_f","date","monthcode","extraction_name","n_riffle","comment")
samples_a$n_riffle[is.na(samples_a$n_riffle)] <- 0 # checked a, and both these samples had two edge samples in them
# samples_a[!samples_a$smpcode_f %in% unq_smpcodes,]  # 3 French I sites, not sampled: exclude them
samples_a <- samples_a[samples_a$smpcode_f %in% unq_smpcodes,]
# MC removed hyphens from extraction names between tables...
samples_a$extraction_name <- gsub("-","",samples_a$extraction_name)
# And 8 still inconsistent....
# inconsistent_smpcodes <- samples_a$smpcode_f[!samples_a$extraction_name %in% miseq_samples$extraction_name]
# unique(miseq_samples$extraction_name[miseq_samples$smpcode_mc %in% inconsistent_smpcodes])
# samples_a$extraction_name[samples_a$smpcode_f %in% inconsistent_smpcodes]
# OK...all but one extraction_name in samples_a missing final digit, and for the last one, the digit is missing in miseq_samples! Fix the exception first
samples_a$extraction_name[samples_a$extraction_name == "S21LER234922"] <- "S21LER23492"
inconsistent_ens <- samples_a$extraction_name[!samples_a$extraction_name %in% miseq_samples$extraction_name]
samples_a$extraction_name[match(inconsistent_ens, samples_a$extraction_name)] <- 
  paste0(inconsistent_ens, right(samples_a$sitecode[match(inconsistent_ens, samples_a$extraction_name)],1))
# sum(!samples_a$extraction_name %in% miseq_samples$extraction_name)  # 0: all match now
# sum(samples_a$monthcode != calcMonthDate(samples_a$date)) # zero: good
# sum(!samples_a$smpcode_f %in% miseq_samples$smpcode) # 0: good - all 137 samples_a$smpcode_f in miseq_samples.
samples_a <- samples_a[samples_a$smpcode_f %in% unq_smpcodes,]
sum(samples_a$sitecode != substr(samples_a$smpcode_f,5,nchar(samples_a$smpcode_f))) # 0 all consistent.
sum(!samples_a$reach %in% sites$reach) # 0
sum(!samples_a$sitecode %in% sites$sitecode) # 14 sitecodes not consistent with sites table
samples_a$sitecode[!samples_a$sitecode %in% sites$sitecode]
# All except STV_2279-4 and UYT-199-4 are reaches rather than sitecodes
samples_a$sitecode[samples_a$sitecode == "UYT-199-4"] <- "UYT-199-3" # smpcode weirdly right
samples_a$sitecode[samples_a$sitecode == "STV-2279-4"] <- "STV-2279-5"
samples_a$smpcode[samples_a$sitecode == "STV-2279-5"] <- "384-STV-2279-5"
# fix the remaining 12
inconsistent_sites <- samples_a$sitecode[!samples_a$sitecode %in% sites$sitecode]
corrected_sites <- sites$sitecode[match(inconsistent_sites,sites$reach)]

samples_a$sitecode[match(inconsistent_sites, samples_a$sitecode)] <- corrected_sites
samples_a$smpcode[match(corrected_sites, samples_a$sitecode)] <- 
  paste(samples_a$monthcode, samples_a$sitecode, sep = "-")[match(corrected_sites, samples_a$sitecode)]
# additional fields to be consistent with c and d below
samples_a$sitecode_mc <- samples_a$sitecode
samples_a <- dplyr::mutate(samples_a, reach = sites$reach[match(samples_a$sitecode, sites$sitecode)],
                           .after = "sitecode")

sum(!samples_a$sitecode %in% sites$sitecode) # 0
sum(!samples_a$extraction_name %in% miseq_samples$extraction_name) # 0 (so sites, samples_a, and miseq_samples, consistent)

b <- as.data.frame(readxl::read_excel("~/uomShare/wergData/DNA_Barcoding/DNA barcoding sample processing Aut2022.xlsx"))
samples_b <- data.frame(sitecode  = b$sitecode, 
                        smpcode_f = NA, # No sample code field in b
                        date = b$`Actual Date_uom`,
                        extraction_name = b$`Extraction name`,
                        n_riffle = b$`number of riffle`)
# 26 NA n_riffles, but checking the 'no_of_kick_jars' field n b, all these samples included a riffle
samples_b$n_riffle <- 1
# remove whitespaces
samples_b$sitecode <- stringr::str_trim(samples_b$sitecode)
samples_b$extraction_name <- stringr::str_trim(samples_b$extraction_name)
samples_b$date[samples_b$sitecode == "TOO-4334-2"] <- as.Date("2022-05-12")
samples_b$date[is.na(samples_b$date)] <- b$`Actual Date_ghd`[is.na(samples_b$date)]
samples_b$monthcode <- calcMonthDate(samples_b$date)
samples_b$comment <- NA
samples_b$comment[!is.na(b$`Actual Date_ghd`)] <- "collected by GHD"
samples_b$comment[!is.na(b$`Actual Date_uom`)] <- "collected by G Hehir, Unimelb"
samples_b$comment[samples_b$sitecode == "TOO-4334-2"] <- "collected by GHD"
sum(!samples_b$extraction_name %in% miseq_samples$extraction_name)  # 0: all consistent
# so, as there is no sample code field in f, make smpcode_f equal to smpcodes in miseq_samples
samples_b$smpcode_f <- miseq_samples$smpcode_mc[match(samples_b$extraction_name, miseq_samples$extraction_name)]
samples_b$smpcode <- samples_b$smpcode_f
# Mix of reaches and sitecodes
samples_b$reach <- samples_b$sitecode
samples_b$reach[stringr::str_count(samples_b$sitecode,"-") == 2] <- 
  substr(samples_b$reach[stringr::str_count(samples_b$sitecode,"-") == 2], 1, 
         nchar(samples_b$reach[stringr::str_count(samples_b$sitecode,"-") == 2]) - 2)
# WM1 (Wombat Creek) with wrong sitecode.
samples_b$reach[samples_b$reach == "WM1-232"] <- "WM1-244"
samples_b$sitecode <- sites$sitecode[match(samples_b$reach, sites$reach)]
samples_b$smpcode <- paste(samples_b$monthcode, samples_b$sitecode, sep = "-")
# additional field to be consistent with c and d below
samples_b$sitecode_mc <- samples_b$sitecode

sum(!samples_b$sitecode %in% sites$sitecode) # 0 sitecodes consistent with sites table
sum(!samples_b$smpcode_f %in% miseq_samples$smpcode) # 0: good - all 137 samples_a$smpcode_f in miseq_samples.
sum(!samples_b$sitecode %in% sites$sitecode) # 0
sum(!samples_b$extraction_name %in% miseq_samples$extraction_name) # 0 (so sites, samples_a, and miseq_samples, consistent)

c <- as.data.frame(readxl::read_excel("~/uomShare/wergData/DNA_Barcoding/DNA barcoding sample processing Spr2022.xlsx"))
samples_c <- data.frame(sitecode  = c$sitecode, 
                        smpcode_f = NA, # No sample code field in c
                        date = c$`Date sampled`,
                        monthcode = NA,
                        extraction_name = c$`Extraction name`,
                        n_riffle = c$`number of riffle`,
                        comment = "collected by G Hehir, Unimelb",
                        reach = c$reach)
# 10 is.na n_riffles, but all included riffles according to 'sampling comments'
samples_c$n_riffle[is.na(samples_c$n_riffle)] <- 1
# remove 6 empty rows
samples_c <- samples_c[!is.na(samples_c$reach),]
samples_c$monthcode <- calcMonthDate(samples_c$date)
# 14 records with no extraction name: not collected, therefore remove
samples_c <- samples_c[!is.na(samples_c$extraction_name),]
sum(!samples_c$extraction_name %in% miseq_samples$extraction_name)  # 1 inconsistent - a typo
samples_c$extraction_name[samples_c$extraction_name == "S22KRY738"] <- "S22KRY758"
# so, as there is no sample code field in c, make smpcode_f equal to smpcodes in miseq_samples
samples_c$smpcode_f <- miseq_samples$smpcode_mc[match(samples_c$extraction_name, miseq_samples$extraction_name)]
samples_c$smpcode <- samples_c$smpcode_f
# These are incorrect, using only reach codes, and these are based on mwstr v1.2 sitecodes...
samples_c$reach <- substr(samples_c$sitecode, 1, nchar(samples_c$sitecode)-1)
samples_c$reach <- gsub("-","_", samples_c$reach)
# all but five of these are in the mwbugs database, and the other 5 are just minor differences
diff_reaches <-  c("CVD_644" = "CVD_663", "EDG_714" = "EDG_720", "LON_611" = "LON_598",
                   "M3D_215" = "M3D_191","TN2_118" = "TN2_104")
samples_c$reach[match(names(diff_reaches), samples_c$reach)] <- diff_reaches
sum(!samples_c$reach %in% sites$reach_v12) #1 inconsistent - problematic french I site
samples_c$reach[samples_c$reach == "TNK_1992"] <- "TNK_2042"

samples_c$sitecode_mc <- samples_c$sitecode
samples_c$sitecode <- sites$sitecode[match(samples_c$reach, sites$reach_v12)]
samples_c$smpcode <- paste(samples_c$monthcode, samples_c$sitecode, sep = "-")

   # 2 sitecodes consistent with sites table
samples_c$sitecode[match(c("396-MUC-987","395-TNK-1992"), samples_c$smpcode_f)] <- c("MUC-987-0","TNK-2042-3")
samples_c$smpcode[match(c("396-MUC-987","395-TNK-1992"), samples_c$smpcode_f)] <- c("396-MUC-987-0","396-TNK-2042-3")

sum(!samples_c$sitecode %in% sites$sitecode) 
sum(!samples_c$smpcode_f %in% miseq_samples$smpcode) # 0: good - all 137 samples_a$smpcode_f in miseq_samples.
sum(!samples_c$extraction_name %in% miseq_samples$extraction_name) # 0 (so sites, samples_a, and miseq_samples, consistent)

d <- as.data.frame(readxl::read_excel("~/uomShare/wergData/DNA_Barcoding/DNA barcoding sample processing Aut2023.xlsx"))
# Just the ten repeated samples - note FER_648f = FER_648c (more or less!)
samples_d <- data.frame(sitecode  = d$sitecode, 
                        smpcode_f = NA, # No sample code field in c
                        date = d$`Date sampled`,
                        monthcode = NA,
                        extraction_name = d$`Extraction name`,
                        comment = "collected by G Hehir, Unimelb",
                        n_riffle = d$`number of riffle`,
                        reach = d$reach)
# 2 without dates collected by GHD - discerned the date from their supplied file 
# "~/uomShare/wergProj/bugDatabase/data_as_supplied/2022-23 MW Macroinvertebrate Data - Board Reporting.xlsx"
samples_d$date[samples_d$sitecode %in% c("YAR_3589h", "FEH_4184d")] <- as.Date("2023-05-10")
samples_d$monthcode <- calcMonthDate(samples_d$date)
samples_d$comment[d$`Sample by` == "GHD"] <- "collected by GHD"
sum(!samples_d$extraction_name %in% miseq_samples$extraction_name)  # 10 all inconsistent!
unique(miseq_samples$extraction_name[grep("A23",miseq_samples$extraction_name)])
# Not even close to consistent! Just use a manual switch
incons_ens <- c("A23BNY29042" = "A23BNY2956",
                "A23DNG318813" = "A23DNG13881",
                "A23DNG79574" = "A23DNG7957",
                "A23FEH41843" = "A23FEH4184",
                "A23FER6475" =  "A23FER648",
                "A23GRD75620" = "A23GRD7570",
                "A23TOO43342" = "A23TOO4269",
                "A23UYT1993" =  "A23UYT199",
                "A23WMI232" = "A23WM1244",
                "A23YAR36377" = "A23YAR3589")
samples_d$extraction_name[match(incons_ens, samples_d$extraction_name)] <- names(incons_ens)
sum(!samples_d$extraction_name %in% miseq_samples$extraction_name)  # 0: consistent now
# so, as there is no sample code field in d, make smpcode_f equal to smpcodes in miseq_samples
samples_d$smpcode_f <- miseq_samples$smpcode_mc[match(samples_d$extraction_name, miseq_samples$extraction_name)]
samples_d$smpcode <- samples_d$smpcode_f
# These are incorrect, using only reach codes, and these are based on mwstr v1.2 sitecodes...
samples_d$reach <- substr(samples_d$sitecode, 1, nchar(samples_d$sitecode)-1)
sum(!samples_d$reach %in% sites$reach_v12) # 0: all in the database
# so, correct sitecodes
samples_d$sitecode_mc <- samples_d$sitecode
samples_d$sitecode <- sites$sitecode[match(samples_d$reach, sites$reach_v12)]
samples_d$smpcode <- paste(samples_d$monthcode, samples_d$sitecode, sep = "-")
# 2 is.na n_riffle - not clear from table d, but these sites included riffles in c, b and a
samples_c$n_riffle[samples_c$sitecode %in% samples_d$sitecode[is.na(samples_d$n_riffle)]]
samples_d$n_riffle[is.na(samples_d$n_riffle)] <- 1

sum(!samples_d$sitecode %in% sites$sitecode) # 0 sitecodes consistent with sites table
sum(!samples_d$smpcode_f %in% miseq_samples$smpcode) # 0: good - all samples_a$smpcode_f in miseq_samples.
sum(!samples_d$extraction_name %in% miseq_samples$extraction_name) # 0 (so sites, samples_a, and miseq_samples, consistent)

samples_miseq <- rbind(samples_a, samples_b[,match(names(samples_a),names(samples_b))],
                     samples_c[,match(names(samples_a),names(samples_c))],
                     samples_d[,match(names(samples_a),names(samples_d))])

# Make samples table ready for import into mwbugs database
new_samples <- as.data.frame(matrix(nrow = nrow(samples_miseq), ncol = ncol(samples)))
names(new_samples) <- names(samples)
for(i in 1:ncol(samples_miseq)){
  if(names(samples_miseq)[i] %in% names(samples)) 
    new_samples[names(samples) == names(samples_miseq)[i]] <- samples_miseq[,i]
}
new_samples$sitecode_v12 <- sites$sitecode_v12[match(new_samples$sitecode, sites$sitecode)]
new_samples$sourcecode <- 66
new_samples$processing_method <- processing_methods$processing_method[processing_methods$pcode == "P"] 
  # i.e. "two samples combined, homogenized, DNA-metabarcoded"
new_samples$collection_method <- "RBA edge-riffle combined"
new_samples$collection_method[samples_miseq$n_riffle == 0] <- "RBA two edges combined"
new_samples$comment = "count value is number of records from 3 miseq replicates"
new_samples$comment[new_samples$smpcode == "384-ANS-1611-6-DP"] <- "count value is number of records from 2 miseq replicates"
new_samples$embargoed <- 1
new_samples$old_samplecode <- samples_miseq$smpcode_f
new_samples$dry_no_sample <- 0  
smpcode_vec <- new_samples$smpcode
new_samples$smpcode <- paste0(smpcode_vec, "-DP")
new_samples$smpcode[samples_miseq$n_riffle == 0] <- paste0(smpcode_vec[samples_miseq$n_riffle == 0], "-EP")

# Add the new sourcecode to the mwbugs spv table.
new_spv <- data.frame(sourcecode = 66, project = "ARC LP DNA metabarcoding", reference = NA,
                      e_source = "github.com/mecarew/metabarcoding_workflow/src_water_prod/Appendix3_import_miseq_to_mwbugs.qmd",
                      file_locations = 
"wergStaff/MelCarew/git-data/metabarcoding_workflow",
lab = "PEARG and WERG Unimelb, led by Melissa Carew",
comment = NA)
```


Assemble biological data for import into database

```{r}
# # some max_p_identity inconsistencies
# x <- unique(bugdata_spp[c("asv_code","max_p_identity","amplicon")])
# x <- x[duplicated(x$asv_code) | duplicated(x$asv_code, fromLast = TRUE),]
# x[order(x$asv_code),]

# Ad hoc corrections to 3 inconsistent max_p_identity values (the first concerning
# "Austropsyche sp. B-AAW3080", above)
bugdata_spp$max_p_identity[bugdata_spp$asv_code == "00f4bc36686affa317cbbdbad7a65991"] <- 99.7
bugdata_spp$max_p_identity[bugdata_spp$asv_code == "bfb774b22881eee84f0f221daeb9d157"] <- 100
bugdata_spp$max_p_identity[bugdata_spp$asv_code == "ed6b0b7d5ab0813489018124e4f46c1e"] <- 99.7

bugdata_spp$extraction_name <- substr(bugdata_spp$sample,1,nchar(bugdata_spp$sample)-4)
bugdata_spp$smpcode <- samples_miseq$smpcode[match(bugdata_spp$extraction_name, 
                                                   samples_miseq$extraction_name)]
# The NAs introduced here are the three samples (1418 records) with smpcodes starting with 337 excluded above
sequences <- unique(bugdata_spp[c("asv_code","asv_seq","max_p_identity","amplicon")])

biota_wrkng <- bugdata_spp[c("sample","smpcode","asv_code","species")]
biota_wrkng <- biota_wrkng[!is.na(biota_wrkng$smpcode),]
biota_wrkng$bugcode <- taxon_all$shortcode[match(biota_wrkng$species, taxon_all$taxon)]
for(i in which(is.na(biota_wrkng$bugcode))){
    biota_wrkng$bugcode[i] <- new_spp$bugcode[new_spp$species == biota_wrkng$species[i]]
   }
sample_asv <- unique(biota_wrkng[c("asv_code","smpcode")])
bugcode_asv <- unique(biota_wrkng[c("asv_code","bugcode")])

biota_agg <- aggregate(biota_wrkng$bugcode, by = list(smpcode = biota_wrkng$smpcode,
                                                      taxoncode = biota_wrkng$bugcode,
                                                      species = biota_wrkng$species), FUN = length)

new_biota <- data.frame(smpcode = biota_agg$smpcode,
                        taxoncode = biota_agg$taxoncode,
                        count = biota_agg$x,
                        taxon = biota_agg$species, 
                        coarsepick = 0, 
                        notes = "",
                        originalbugcode = NA,
                        shortcode = NA)

# Add records only identified to genus:
bugdata_gen$extraction_name <- substr(bugdata_gen$sample,1,nchar(bugdata_gen$sample)-4)
bugdata_gen$smpcode <- samples_miseq$smpcode[match(bugdata_gen$extraction_name, 
                                                   samples_miseq$extraction_name)]
sequences <- rbind(sequences,
                   unique(bugdata_gen[c("asv_code","asv_seq","max_p_identity","amplicon")]))
# NAs introduced as above
biota_wrkng <- bugdata_gen[c("sample","smpcode","asv_code","genus")]
biota_wrkng <- biota_wrkng[!is.na(biota_wrkng$smpcode),]
biota_wrkng$bugcode <- paste0(taxon_gen$gencode[match(biota_wrkng$genus, taxon_gen$genus)],"99")

biota_agg <- aggregate(biota_wrkng$bugcode, by = list(smpcode = biota_wrkng$smpcode,
                                                      taxoncode = biota_wrkng$bugcode,
                                                      species = biota_wrkng$genus), 
                                                      FUN = length)
biota_agg$species <- paste0(biota_agg$species, " sp. (Unident.)")

new_biota <- rbind(new_biota, data.frame(smpcode = biota_agg$smpcode,
                        taxoncode = biota_agg$taxoncode,
                        count = biota_agg$x,
                        taxon = biota_agg$species, 
                        coarsepick = 0, 
                        notes = "",
                        originalbugcode = NA,
                        shortcode = NA))

sample_asv <- rbind(sample_asv, unique(biota_wrkng[c("asv_code","smpcode")]))
bugcode_asv <- rbind(bugcode_asv, unique(biota_wrkng[c("asv_code","bugcode")]))

# Add records only identified to family:
bugdata_fam$extraction_name <- substr(bugdata_fam$sample,1,nchar(bugdata_fam$sample)-4)
bugdata_fam$smpcode <- samples_miseq$smpcode[match(bugdata_fam$extraction_name, 
                                                   samples_miseq$extraction_name)]
# NAs introduced as above
sequences <- rbind(sequences,
                   unique(bugdata_fam[c("asv_code","asv_seq","max_p_identity","amplicon")]))
biota_wrkng <- bugdata_fam[c("sample","smpcode","asv_code","family")]
biota_wrkng$extraction_name <- substr(biota_wrkng$sample,1,nchar(biota_wrkng$sample)-4)
biota_wrkng$smpcode <- samples_miseq$smpcode[match(biota_wrkng$extraction_name, 
                                                   samples_miseq$extraction_name)]
biota_wrkng <- biota_wrkng[!is.na(biota_wrkng$smpcode),]
biota_wrkng$bugcode <- paste0(taxon_fam$famcode[match(biota_wrkng$family, taxon_fam$family)],"9999")
biota_agg <- aggregate(biota_wrkng$bugcode, by = list(smpcode = biota_wrkng$smpcode,
                                                      taxoncode = biota_wrkng$bugcode,
                                                      species = biota_wrkng$family), 
                                                      FUN = length)
biota_agg$species <- paste0(biota_agg$species, " sp. (Unident.)")
new_biota <- rbind(new_biota, data.frame(smpcode = biota_agg$smpcode,
                        taxoncode = biota_agg$taxoncode,
                        count = 1,
                        taxon = biota_agg$species, 
                        coarsepick = 0, 
                        notes = "",
                        originalbugcode = NA,
                        shortcode = NA))

sample_asv <- rbind(sample_asv, unique(biota_wrkng[c("asv_code","smpcode")]))
bugcode_asv <- rbind(bugcode_asv, unique(biota_wrkng[c("asv_code","bugcode")]))

#####
x <- bugcode_asv[duplicated(bugcode_asv$asv_code) | duplicated(bugcode_asv$asv_code, fromLast = TRUE),]
# 0 asv_codes with > 1 identification

new_biota <- new_biota[order(new_biota$smpcode),]
```


```{r}
bins <- as.data.frame(readxl::read_excel("data/asv_library_corrected.xlsx", sheet = 3))
bins <- bins[!is.na(bins$bin_uri),]

miseq_data <- rbind(bugdata_spp[c("sample","asv_code","reads")],
                    bugdata_gen[c("sample","asv_code","reads")],
                    bugdata_fam[c("sample","asv_code","reads")])
miseq_data <- dplyr::mutate(miseq_data, 
                            smpcode = miseq_data$smpcode[match(miseq_data$sample, 
                                                               miseq_samples$seq_code)],
                            .after = sample)

 reads_tab <- dplyr::mutate(miseq_data, 
                            smpcode = miseq_samples$smpcode[match(miseq_data$sample, miseq_samples$seq_code)],
                            bugcode = bugcode_asv$bugcode[match(miseq_data$asv_code, bugcode_asv$asv_code)], 
                            .after = sample)
 reads_tab <- dplyr::mutate(reads_tab, 
                            taxon = new_biota$taxon[match(paste0(reads_tab$smpcode,reads_tab$bugcode),
                                                          paste0(new_biota$smpcode,new_biota$taxoncode))],
                            .after = bugcode)
```

```{r}
#Build xlsx file for inspection prior to adding these data to the mwbugs database
table_list <- list(sites = sites[sites$sitecode %in% unique(new_samples$sitecode),],
                   new_samples = new_samples,
                   new_biota = new_biota,
                   collection_methods = collection_methods[collection_methods$ccode %in% c("D","E"),],
                   processing_methods = processing_methods[processing_methods$pcode %in% c("P"),],
                   new_spv = new_spv,
                   sequences = sequences,
                   bugcode_asv = bugcode_asv, 
                   sample_asv = sample_asv,
                   new_taxon_spp = new_taxon_spp,
                   new_morphospp_etc = new_morphospp_etc, 
                   miseq_samples = miseq_samples,
                   miseq_data = miseq_data,
                   bins = bins)
WriteXLS::WriteXLS(table_list, "~/temp/arc_metabarcoding_data_for_mwbugs_db.xlsx")

```
