---
title: "R Notebook"
output: html_notebook
---


Scripts for manipulating data from final long format


#convert long form back to full data summary with data filled and filtered
Note: This code will need to be configured for each dataset, so that the rows containing the read counts are specified (e.g. pivot_longer(cols = X10BNY29042rep1:Cont3rep1), the 'X10BNY29042rep1:Cont3rep1' will need to be changed to the first and last column name
Miseq13:X10BNY29042rep1:Cont3rep1 
Miseq15:X10FEH41841rep3:S21YAR36377rep3
Miseq16:Contrep1:S21WAY30960rep3
Miseq17:A22BNY29042rep1:S21YZJ1175rep6
Miseq18:A22ALD26081rep1:S22YZ8202rep3
Miseq19:A22BF3373rep1:S22YPB224rep3
Miseq20:A23BNY29042rep1:S22SPL263rep6

```{r}
# Step 1: Replace empty cells with 0 before pivoting
final_data_filled_clean <- filtered_dataframe %>%
  mutate(across(, ~ replace_na(.x, 0)))  # Replace NA with 0 in specified columns

# Step 2: Pivot longer
metab_long <- final_data_filled_clean %>%
  pivot_longer(
    cols = Contrep1:S21WAY30960rep3,  # Specify the range or individual column names
    names_to = "sample",
    values_to = "value"
  ) %>%
  # Ensure that the 'value' column is numeric and any empty strings are replaced with 0
  mutate(value = as.numeric(value)) %>%
  filter(value != 0) %>%  # Optionally filter out rows where value is zero
  select(-matches("^X(\\.\\d+)?$"))  # Remove columns named X, X.1, ..., X.11

# Step 3: Pivot wider to transform back to wide format
metab_wide <- metab_long %>%
  pivot_wider(
    names_from = sample,
    values_from = value,
    values_fill = list(value = 0)  # Ensure missing values in the new wide format are filled with 0
  )

# List of columns to modify
cols_to_modify <- c('kingdom', 'phylum', 'class', 'order', 'family', 'genus', 'species')

# Convert specified columns to character type
metab_wide[cols_to_modify] <- lapply(metab_wide[cols_to_modify], as.character)

# Replace NA values with empty strings in specified columns
metab_wide[cols_to_modify][is.na(metab_wide[cols_to_modify])] <- ""

# Remove 'X' at the start of column headings
colnames(metab_wide) <- sub("^X", "", colnames(metab_wide))

new_order <- c(setdiff(names(metab_wide), c("asv_seq", "amplicon")), "asv_seq", "amplicon")

# Reorder the columns
metab_wide <- metab_wide[, new_order]


# Print the first few rows of the resulting dataframe
print(head(metab_wide))
```

# save the filtered filled full dataframe with the PCR replcate seperate
```{r}
# Generate a timestamp
timestamp <- format(Sys.time(), "%d_%m_%Y")

# Create the file name with the timestamp (set Miseq run number)
file_name <- paste0(uomshare, "/final_data_summaries/", miseq_folder, "_vsearch_data_summary_filled_filt_PCR_", timestamp, ".csv")
# Save the dataframe with the new file name
write.csv(metab_wide, file = file_name, row.names = FALSE)
```

## There are a number of options for exporting data for further analysis

# OPTION 1: Summarize values by site (with PCR replicates combined)
```{r}
# Summarize the data
summarized_df <- filtered_dataframe %>%
  group_by(site, kingdom, phylum, class, order, family, genus, species, max_p_identity) %>%  # Group by specified columns including max_p_identity
  summarise(reads = sum(reads, na.rm = TRUE), .groups = 'drop') %>%  # Sum the reads for each group
  group_by(site, kingdom, phylum, class, order, family, genus, species) %>%  # Group by the main columns again
  summarise(
    reads = sum(reads, na.rm = TRUE),  # Sum the reads for each group
    max_p_identity = ifelse(all(is.na(max_p_identity)), NA, max(max_p_identity, na.rm = TRUE)),  # Handle all NA case
    .groups = 'drop'
  )

# Collapse rows by kingdom, phylum, class, order, family, genus, species
collapsed_df <- summarized_df %>%
  group_by(site, kingdom, phylum, class, order, family, genus, species) %>%
  summarise(
    reads = sum(reads, na.rm = TRUE),  # Sum the reads for each group
    max_p_identity = ifelse(all(is.na(max_p_identity)), NA, max(max_p_identity, na.rm = TRUE)),  # Get max value of max_p_identity
    .groups = 'drop'
  )

# Pivot wider to transform the site names into columns
wide_summarised_df <- collapsed_df %>%
 pivot_wider(names_from = site, values_from = reads, values_fill = 0)  # Fill missing values with 0

# List of columns to modify
cols_to_modify <- c('kingdom', 'phylum', 'class', 'order', 'family', 'genus', 'species')

# Convert specified columns to character type
wide_summarised_df[cols_to_modify] <- lapply(wide_summarised_df[cols_to_modify], as.character)

# Replace NA values with empty strings in specified columns
wide_summarised_df[cols_to_modify][is.na(wide_summarised_df[cols_to_modify])] <- ""

#subset out controls
controls <- wide_summarised_df %>% 
  select(kingdom, phylum, class, order, family, genus, species, max_p_identity, DrosX, MAR2T, Cont)
controls <- controls %>%
 filter(MAR2T != 0 | DrosX != 0, Cont != 0)
filec <- paste0(uomshare, "/final_data_summaries/", miseq_folder, "_controls_", timestamp, ".csv")
 write.csv(controls, file = filec, row.names = FALSE)

wide_summarised_df <- wide_summarised_df %>% 
  select(-DrosX, -MAR2T, -Cont)

# Print the resulting wide summarized data frame
print(wide_summarised_df)

```

#Now aggregate species data so that the same species identification do not occur multiple times
```{r}
library(dplyr)

# List of columns to modify
cols_to_modify <- c('kingdom', 'phylum', 'class', 'order', 'family', 'genus', 'species')

# Convert specified columns to character type
wide_summarised_df[cols_to_modify] <- lapply(wide_summarised_df[cols_to_modify], as.character)

# Replace NA values with empty strings in specified columns
wide_summarised_df[cols_to_modify][is.na(wide_summarised_df[cols_to_modify])] <- ""

# Print the resulting wide summarized data frame
print(wide_summarised_df)

# List of columns to group by
group_cols <- c('kingdom', 'phylum', 'class', 'order', 'family', 'genus', 'species')

# Identify the columns that are site_per (excluding group columns and max_p_identity)
site_cols <- setdiff(colnames(wide_summarised_df), c(group_cols, 'max_p_identity'))

# Collapse rows by kingdom, phylum, class, order, family, genus, species
collapsed_wide_summarised_df <- wide_summarised_df %>%
  group_by(across(all_of(group_cols))) %>%
  summarise(
    max_p_identity = max(max_p_identity, na.rm = TRUE),  # Get max value of max_p_identity
    across(all_of(site_cols), function(x) sum(x, na.rm = TRUE)),  # Sum the values for the site_per columns
    .groups = 'drop'
  )

# Print the collapsed data frame
print(collapsed_wide_summarised_df)
```

# save file
```{r}
# save filled dataframe
timestamp <- format(Sys.time(), "%d_%m_%Y")

# Create the file name with the timestamp (NOTE:Change the MiSeq number to the dataset you are working with)
file_name1 <- paste0(uomshare, "/final_data_summaries/", miseq_folder, "_final_data_summary_", timestamp, ".csv")
 write.csv(wide_summarised_df, file = file_name1, row.names = FALSE)
```