---
title: "Appendix 3. Importing miseq data into Melbourne Water macroinvertebrate database"
format: html
editor_options: 
  chunk_output_type: console
---

## Introduction

This document reads the miSeq data prepared in the first 6 metabarcoding workflow documents. The code in this document consolidates the data, and formats them ready to be imported into the [Melbourne Water macroinvertebrate database](https://tools.thewerg.unimelb.edu.au/mwbugs/) (mwbugs database).  The code in the document begins by making a connection to the main PostgreSQL database on the University of Melbourne Server. For users outside the unimelb network, the database can be accessed using the instructions in Appendix B of Walsh (2023)[^1].

[^1]: Walsh, C.J. (2023) The Melbourne Water macroinvertebrate database: user’s manual. Version 2. Melbourne Waterway Research-Practice Partnership Report 20.1a. School of Agriculture, Forest and Ecosystem Sciences, The University of Melbourne, Melbourne. URL: https://tools.thewerg.unimelb.edu.au/mwbugs_manual/.

The document also contains code for identifying new species not yet included in the mwbugs database taxonomy tables, and determining new bugcodes for those species.  When this document was first prepared xxx new species were identified and added to the database.  The code is retained here for future use, when further new species are encountered.

```{r}
#| label: setup

source("https://tools.thewerg.unimelb.edu.au/data/mwbugs/bug_database_functions.R")

# Connect to the mwbugs database and load all tables (using the load_all_mwbugs_tables.R 
#  from bug_database_functions.R)

db <- RPostgres::dbConnect(drv = RPostgres::Postgres(), dbname = "mwbugs")

load_all_mwbugs_tables(db)

# Function equivalent to excel's =RIGHT() and =LEFT() functions
right <- function(x, n) {substr(x, nchar(x) - n + 1, nchar(x))}
left <- function(x, n) {substr(x, 1, nchar(x) - n)}

# List of candidate character pairs for filling bugcodes
# avoid confusion of 0s and Os - exclude Os, and exclude Zs which are reserved for unknowns
all_prs <- expand.grid(x = c(0:9,LETTERS[-c(15,26)]), 
                       y = c(0:9,LETTERS[-c(15,26)]), stringsAsFactors = FALSE)
all_prs <- all_prs[order(all_prs$x),]
all_prs$prs <- paste0(all_prs$x, all_prs$y)
# and remove all 0s combined with letters and 00
zeros <- which((grepl("0", all_prs$prs) & 
                  grepl("[A-Z]", all_prs$prs)) | all_prs$pr == "00")
all_prs <- all_prs[-zeros,]
# order with numbers first, and mark numbers
all_prs$number <- 0
all_prs$number[!is.na(as.numeric(all_prs$prs))] <- 1
all_prs <- all_prs[order(all_prs$number, decreasing = TRUE),]
# order alphabetic codes so that A1 comes before 1A etc...(probably won't use 1A etc)
all_prs <- rbind(all_prs[all_prs$number == 1,],
                 all_prs[grepl("[A-Z]", all_prs$x) & all_prs$number == 0,],
                 all_prs[!grepl("[A-Z]", all_prs$x) & all_prs$number == 0,])

```

The following chunk compiles the miseq_samples, samples, and bugdata tables from the source files for 6 miseq runs, checking for and correcting consistency between the tables (with checks back to lab notes for corrections.)

The bugdata is separated into four tables (spp, gen, fam, other) to separate records identified to species (or species group) from those identified to genus, family or a higher level of taxonomy (other).

```{r}
#| echo: true

# Compile all sample data
miseq_dirs <- c(15,16,17,18,19,20)

for(i in 1:length(miseq_dirs)){
  miseq_path <- paste0("~/uomShare/wergStaff/MelCarew/git-data/metabarcoding_workflow/miseq", miseq_dirs[i],"/")
  miseq_sample_info <- paste0("miseq",miseq_dirs[i], "_sample_information.xlsx")
  samples_i <- rio::import_list(paste0(miseq_path,miseq_sample_info))
  miseq_samples_i <- samples_i$miseq_samples
  miseq_samples_i$miseq_run <- miseq_dirs[i]
  #@@ table naming error in miseq15
  names(samples_i)[names(samples_i) == "sample"] <- "samples"
  smpcodes_i <- samples_i$samples
  #@@ column inconsistencies in miseq15, 18, 19, 20
  names(miseq_samples_i)[names(miseq_samples_i) == "comments"] <- "comment"
  names(miseq_samples_i)[names(miseq_samples_i) == "notes"] <- "comment"
  smpcodes_i <- smpcodes_i[,!names(smpcodes_i) %in% c("...7","amplicon")]
  names(miseq_samples_i)[names(miseq_samples_i) == "position"] <- "plate_position"
  names(miseq_samples_i)[names(miseq_samples_i) == "i7 index"] <- "i7_index"
  names(miseq_samples_i)[names(miseq_samples_i) == "i5 index"] <- "i5_index"
  
#  bugdata_file <- dir(paste0(miseq_path, "metabarcoding_workflow/final_modeling_data"))
#  Most recent file - check contents of bugdata_file above to confirm!
  bugdata_file <- paste0("miseq", miseq_dirs[i], 
                         # MC revised miseq18 data after first run of this file identified a problem
                         ifelse(miseq_dirs[i] == 18, 
                                "_final_data_long_format_26_02_2025.csv",
                                "_final_data_long_format_18_02_2025.csv"))
  bugdata_i <- read.csv(paste0(miseq_path, "metabarcoding_workflow/final_modeling_data/", bugdata_file))
  #@@ smpcode inconsistencies in miseq18
  miseq_samples_i$smpcode[miseq_samples_i$smpcode == "388-BF3-373-3"] <- "388-BF3-373"
  #@@ seq_code inconsistencies in miseq15, 18
  miseq_samples_i$seq_code <- gsub("30DNG79574","30DNG318813",miseq_samples_i$seq_code)
  miseq_samples_i$seq_code <- gsub("A22XZZ3644","A22X223644",miseq_samples_i$seq_code)
  
#  bugdata_i$sample <- gsub("A22YZ2161","A22YZ2162", bugdata_i$sample)
  report_i <- data.frame(miseq = miseq_dirs[i], no_bug_rows = nrow(bugdata_i), 
                         no_bug_rows_NA = sum(is.na(bugdata_i$species)), no_new_seqs = NA)
  bugdata_i <- only_aqu_inverts(bugdata_i, db)$include
  
  # Check uniqueness and consistency
  if(sum(duplicated(smpcodes_i$smpcode)) > 0)  stop("1")
  if(sum(duplicated(miseq_samples_i$seq_code)) > 0)  stop("2")
  # Check only entries in smpcodes_i table with NA smpcodes are controls/refs
  # Note tolower() necessary because of inconsistencies with capitalization
  if(sum(!tolower(substr(smpcodes_i$seq_code[is.na(smpcodes_i$smpcode)], 1, 4)) %in% 
          tolower(c("MAR2","Dros","Cont","MART","CO1C","COIC"))) > 0) stop("3")
  smpcodes_i <- smpcodes_i[!is.na(smpcodes_i$smpcode) & 
                             !tolower(substr(smpcodes_i$smpcode,1,4)) %in%
                              tolower(c("MAR2","Dros","cont","MART","CO1C","COIC")),]
  miseq_samples_i <- miseq_samples_i[!tolower(substr(miseq_samples_i$seq_code, 1, 4)) 
                                     %in% tolower(c("MAR2","Dros","Cont","MART","CO1C","COIC")),] 
  #@@ smpcodes missing from reps 4,5,and 6 in miseq15
  missing_smpcodes <- which(is.na(miseq_samples_i$smpcode))
  for(j in missing_smpcodes){
    miseq_samples_i$smpcode[j] <- miseq_samples_i$smpcode[miseq_samples_i$seq_code == 
                                                            paste0(left(miseq_samples_i$seq_code[j],4), "rep1")]
  }
  # #####@@@@ remove miseq18 samples that were re-run in miseq19
    if(miseq_dirs[i] == 18){
  smpcodes_i <- smpcodes_i[smpcodes_i$smpcode != miseq_samples_i$smpcode[miseq_samples_i$seq_code == "A22YZ2162rep3"],]
  miseq_samples_i <- miseq_samples_i[!miseq_samples_i$seq_code %in% c("A22YZ2162rep3","A22YZ2162rep1","A22YZ2162rep2"),]
  bugdata_i <- bugdata_i[!bugdata_i$sample %in% c("A22YZ2162rep3","A22YZ2162rep1","A22YZ2162rep2"),]
}
  # ####@@@@ and non-matching sample codes in the bugdata file
  if(miseq_dirs[i] == 19){
    # Miscoding errors corrected as per emails from Mel 4/12/2024 and 10/12/2024
    miseq_samples_i$smpcode[miseq_samples_i$seq_code == "A22YZ7161rep1"] <- "389-YZ2-162"
    miseq_samples_i$seq_code[miseq_samples_i$seq_code == "A22YZ7161rep1"] <- "A22YZ2162rep1"
    bugdata_i$sample[bugdata_i$sample == "A22BF3373rep1"] <- "A22BF3373rep4"
    bugdata_i$sample[bugdata_i$sample == "A22BF3373rep2"] <- "A22BF3373rep5"
    bugdata_i$sample[bugdata_i$sample == "A22BF3373rep3"] <- "A22BF3373rep6"
  }
  if(miseq_dirs[i] == 20){
    # Miscoding errors corrected as per email from Mel 4/12/2024
    bugdata_i$sample[bugdata_i$sample == "S22ALL58rep2"] <- "S22ALL558rep4"
    bugdata_i$sample[bugdata_i$sample == "S22ALL58rep3"] <- "S22ALL558rep5"
    bugdata_i$sample[bugdata_i$sample == "S22ALL58rep4"] <- "S22ALL558rep6"
  }
  ####@@@@
  if(sum(is.na(match(miseq_samples_i$smpcode,smpcodes_i$smpcode))) > 0) stop("4")
  if(sum(!smpcodes_i$smpcode %in% miseq_samples_i$smpcode) > 0) stop("5")

  # Check and compile bug data
  # Consistency between miseq_samples table and bugdata table.
  # MISSING bugdata: rep2 for this smpcode did not produce enough reads and was 
  # not re-run (see email from Mel 4/12/2024). Delete miseq_sample
  miseq_samples_i <- miseq_samples_i[miseq_samples_i$seq_code != "S21ANS16116rep2",]
  # delete re-run samples from 10-site study (see email from Mel 4/12/2024)
  miseq_samples_i <- miseq_samples_i[!miseq_samples_i$seq_code %in% 
      c("Elut30DNG318813rep1","S21Elut30DNG318813rep2","Elut30DNG318813rep3"),]
  smpcodes_i <- smpcodes_i[smpcodes_i$smpcode != "DNG-7957-4",]
  bugdata_i <- bugdata_i[!bugdata_i$sample %in% 
                           c("Elut30DNG79574rep1", "Elut30DNG79574rep2", "Elut30DNG79574rep3"),]
  ##@@@ Incorrect sitecode in miseq18 and 19 and 20
  miseq_samples_i$seq_code <- gsub("X223644","XZZ3644", miseq_samples_i$seq_code)
  bugdata_i$sample <- gsub("X223644","XZZ3644", bugdata_i$sample)
 # bugdata_i$sample <- gsub("ALL58","ALL558", bugdata_i$sample)
  #@@ spurious Xs at start of sample codes in miseq15
  bugdata_i$sample[substr(bugdata_i$sample,1,1) == "X"] <- 
    substr(bugdata_i$sample[substr(bugdata_i$sample,1,1) == "X"], 2,
           nchar(bugdata_i$sample[substr(bugdata_i$sample,1,1) == "X"]))
  if(sum(!unique(bugdata_i$sample) %in% miseq_samples_i$seq_code) > 0) stop("7")
  # if(sum(!miseq_samples_i$seq_code %in% unique(bugdata_i$sample)) > 0) 
  #   cat(miseq_samples_i$seq_code[!miseq_samples_i$seq_code %in% unique(bugdata_i$sample)],
  #       "miseq",miseq_dirs[i],"\n")
  # The above commented-out bit is ok - it identifies a couple of seq_codes that were re-run
  # and a couple of other problems identified in the next chunk.
  
  # unique(bugdata_i$sample)[!unique(bugdata_i$sample) %in% miseq_samples_i$seq_code]
  # miseq_samples_i$seq_code[!miseq_samples_i$seq_code %in% unique(bugdata_i$sample)]
  # As directed by Mel in email of 4 Dec 2024, remove species with MC_ARCO in their name
  bugdata_i$species[grep("MC-ARCO", bugdata_i$species)] <- ""
  # Remove "environmental sample" records 
  bugdata_i <- bugdata_i[!grepl("environmental", bugdata_i$species) & 
                           !grepl("uncultured zooplankton", bugdata_i$species),]
  # ensure no trailing or leading spaces
  bugdata_i$genus <- stringr::str_trim(bugdata_i$genus)
  # Check consistency with sample tables

  bugdata_i_spp <- bugdata_i[bugdata_i$species != "",]
  # Should all be similarity >= 0.97
  if(sum(bugdata_i_spp$max_p_identity < 0.97) > 0) stop("A")
  bugdata_i_gen <- bugdata_i[bugdata_i$species == "" & bugdata_i$genus != "",]
  bugdata_i_fam <- bugdata_i[bugdata_i$species == "" & bugdata_i$genus == "" & bugdata_i$family != "",]
  bugdata_i_oth <- bugdata_i[bugdata_i$species == "" & bugdata_i$genus == "" & bugdata_i$family == "",]
  # These four tables should contain all records from bugdata_i 
  if(sum(nrow(bugdata_i_spp),nrow(bugdata_i_gen),nrow(bugdata_i_fam),nrow(bugdata_i_oth)) != nrow(bugdata_i)) stop("B")
  if(i == 1){
    bugdata_spp <- bugdata_i_spp
    bugdata_gen <- bugdata_i_gen
    bugdata_fam <- bugdata_i_fam
    bugdata_oth <- bugdata_i_oth
    smpcodes <- smpcodes_i
    miseq_samples <- miseq_samples_i
  }else{
    bugdata_spp <- rbind(bugdata_spp, bugdata_i_spp)
    bugdata_gen <- rbind(bugdata_gen, bugdata_i_gen)
    bugdata_fam <- rbind(bugdata_fam, bugdata_i_fam)
    bugdata_oth <- rbind(bugdata_oth, bugdata_i_oth)
    smpcodes <- rbind(smpcodes, smpcodes_i)
    if(sum(miseq_samples_i$seq_code %in% miseq_samples$seq_code) > 0 )  stop("8")
    miseq_samples <- rbind(miseq_samples, miseq_samples_i)
  }
}
# A rogue terrestrial to be deleted.
bugdata_gen <- bugdata_gen[bugdata_gen$genus != "Chrysophtharta",]
```

Some samples were re-run because of insufficient reads (or other reasons), often in a subsequent miseq run.  The following chunk finds such cases (where there are replicate 1, 2, and 3 and 4, 5, and 6), and keeps only replicates 4, 5 and 6.

```{r}

sample_reps <- aggregate(miseq_samples$seq_code, 
                         by = list(seq_code = miseq_samples$smpcode), FUN = length)
# sum(!sample_reps$x[i] %in% c(3,6) > 0)  # 0 as expected
  for(j in which(sample_reps$x == 6)){
    miseq_samples <- miseq_samples[!(miseq_samples$smpcode == sample_reps$seq_code[j] &
                                           right(miseq_samples$seq_code, 1) %in% 1:3),]
  }

bugdata_spp <- bugdata_spp[bugdata_spp$sample %in% miseq_samples$seq_code,]
bugdata_gen <- bugdata_gen[bugdata_gen$sample %in% miseq_samples$seq_code,]
bugdata_fam <- bugdata_fam[bugdata_fam$sample %in% miseq_samples$seq_code,]
bugdata_oth <- bugdata_oth[bugdata_oth$sample %in% miseq_samples$seq_code,]

#### Need to remove these seq_codes from the miseq_sample table  
#  (seq_code A22BT2373rep3 & A23YAR36377rep3 not in bugdata_spp table)
#  miseq_samples[!miseq_samples$seq_code %in% unique(bugdata_spp$sample),]
miseq_samples <- miseq_samples[miseq_samples$seq_code %in% unique(bugdata_spp$sample),]

unq_smpcodes <- unique(miseq_samples$smpcode)
# need to associate correct smpcodes with each miseq_samples record. Keep (commonly) incorrect smpcodes as smpcode_mc
miseq_samples <- dplyr::mutate(miseq_samples, smpcode_mc = smpcode, .after = smpcode)
miseq_samples <- dplyr::mutate(miseq_samples, 
                               extraction_name = substr(miseq_samples$seq_code, 1, nchar(miseq_samples$seq_code) - 4), 
                                .after = smpcode_mc)

# Field/planning sample details
a = as.data.frame(readxl::read_excel("~/uomShare/wergData/DNA_Barcoding/DNA barcoding samples processing Spr2021.xlsx"))
samples_a <- a[c("sitecode","sample_code","Actual Date","monthcode","Extraction name","number of riffle")]
samples_a$comment <- NA
samples_a$comment[!is.na(samples_a$`Actual Date_ghd`)] <- "collected by GHD"
samples_a$comment[!is.na(samples_a$`Actual Date_UoM`)] <- "collected by G Hehir, Unimelb"
samples_a$comment[!(is.na(samples_a$`Actual Date_ghd`) & is.na(samples_a$`Actual Date_uom`))] <- "collector unspecified"
names(samples_a) <- c("sitecode","smpcode_f","date","monthcode","extraction_name","n_riffle","comment")
samples_a$n_riffle[is.na(samples_a$n_riffle)] <- 0 # checked a, and both these samples had two edge samples in them
# samples_a[!samples_a$smpcode_f %in% unq_smpcodes,]  # 3 French I sites, not sampled: exclude them
samples_a <- samples_a[samples_a$smpcode_f %in% unq_smpcodes,]
# MC removed hyphens from extraction names between tables...
samples_a$extraction_name <- gsub("-","",samples_a$extraction_name)
# And 8 still inconsistent....
# inconsistent_smpcodes <- samples_a$smpcode_f[!samples_a$extraction_name %in% miseq_samples$extraction_name]
# unique(miseq_samples$extraction_name[miseq_samples$smpcode_mc %in% inconsistent_smpcodes])
# samples_a$extraction_name[samples_a$smpcode_f %in% inconsistent_smpcodes]
# OK...all but one extraction_name in samples_a missing final digit, and for the last one, the digit is missing in miseq_samples! Fix the exception first
samples_a$extraction_name[samples_a$extraction_name == "S21LER234922"] <- "S21LER23492"
inconsistent_ens <- samples_a$extraction_name[!samples_a$extraction_name %in% miseq_samples$extraction_name]
samples_a$extraction_name[match(inconsistent_ens, samples_a$extraction_name)] <- 
  paste0(inconsistent_ens, right(samples_a$sitecode[match(inconsistent_ens, samples_a$extraction_name)],1))
# sum(!samples_a$extraction_name %in% miseq_samples$extraction_name)  # 0: all match now
# sum(samples_a$monthcode != calcMonthDate(samples_a$date)) # zero: good
# sum(!samples_a$smpcode_f %in% miseq_samples$smpcode) # 0: good - all 137 samples_a$smpcode_f in miseq_samples.
samples_a <- samples_a[samples_a$smpcode_f %in% unq_smpcodes,]
sum(samples_a$sitecode != substr(samples_a$smpcode_f,5,nchar(samples_a$smpcode_f))) # 0 all consistent.
sum(!samples_a$reach %in% sites$reach) # 0
sum(!samples_a$sitecode %in% sites$sitecode) # 14 sitecodes not consistent with sites table
samples_a$sitecode[!samples_a$sitecode %in% sites$sitecode]
# All except STV_2279-4 and UYT-199-4 are reaches rather than sitecodes
samples_a$sitecode[samples_a$sitecode == "UYT-199-4"] <- "UYT-199-3" # smpcode weirdly right
samples_a$sitecode[samples_a$sitecode == "STV-2279-4"] <- "STV-2279-5"
samples_a$smpcode[samples_a$sitecode == "STV-2279-5"] <- "384-STV-2279-5"
# fix the remaining 12
inconsistent_sites <- samples_a$sitecode[!samples_a$sitecode %in% sites$sitecode]
corrected_sites <- sites$sitecode[match(inconsistent_sites,sites$reach)]

samples_a$sitecode[match(inconsistent_sites, samples_a$sitecode)] <- corrected_sites
samples_a$smpcode[match(corrected_sites, samples_a$sitecode)] <- 
  paste(samples_a$monthcode, samples_a$sitecode, sep = "-")[match(corrected_sites, samples_a$sitecode)]
# additional fields to be consistent with c and d below
samples_a$sitecode_mc <- samples_a$sitecode
samples_a <- dplyr::mutate(samples_a, reach = sites$reach[match(samples_a$sitecode, sites$sitecode)],
                           .after = "sitecode")

sum(!samples_a$sitecode %in% sites$sitecode) # 0
sum(!samples_a$extraction_name %in% miseq_samples$extraction_name) # 0 (so sites, samples_a, and miseq_samples, consistent)

b <- as.data.frame(readxl::read_excel("~/uomShare/wergData/DNA_Barcoding/DNA barcoding sample processing Aut2022.xlsx"))
samples_b <- data.frame(sitecode  = b$sitecode, 
                        smpcode_f = NA, # No sample code field in b
                        date = b$`Actual Date_uom`,
                        extraction_name = b$`Extraction name`,
                        n_riffle = b$`number of riffle`)
# 26 NA n_riffles, but checking the 'no_of_kick_jars' field n b, all these samples included a riffle
samples_b$n_riffle <- 1
# remove whitespaces
samples_b$sitecode <- stringr::str_trim(samples_b$sitecode)
samples_b$extraction_name <- stringr::str_trim(samples_b$extraction_name)
samples_b$date[samples_b$sitecode == "TOO-4334-2"] <- as.Date("2022-05-12")
samples_b$date[is.na(samples_b$date)] <- b$`Actual Date_ghd`[is.na(samples_b$date)]
samples_b$monthcode <- calcMonthDate(samples_b$date)
samples_b$comment <- NA
samples_b$comment[!is.na(b$`Actual Date_ghd`)] <- "collected by GHD"
samples_b$comment[!is.na(b$`Actual Date_uom`)] <- "collected by G Hehir, Unimelb"
samples_b$comment[samples_b$sitecode == "TOO-4334-2"] <- "collected by GHD"
sum(!samples_b$extraction_name %in% miseq_samples$extraction_name)  # 0: all consistent
# so, as there is no sample code field in f, make smpcode_f equal to smpcodes in miseq_samples
samples_b$smpcode_f <- miseq_samples$smpcode_mc[match(samples_b$extraction_name, miseq_samples$extraction_name)]
samples_b$smpcode <- samples_b$smpcode_f
# Mix of reaches and sitecodes
samples_b$reach <- samples_b$sitecode
samples_b$reach[stringr::str_count(samples_b$sitecode,"-") == 2] <- 
  substr(samples_b$reach[stringr::str_count(samples_b$sitecode,"-") == 2], 1, 
         nchar(samples_b$reach[stringr::str_count(samples_b$sitecode,"-") == 2]) - 2)
# WM1 (Wombat Creek) with wrong sitecode.
samples_b$reach[samples_b$reach == "WM1-232"] <- "WM1-244"
samples_b$sitecode <- sites$sitecode[match(samples_b$reach, sites$reach)]
samples_b$smpcode <- paste(samples_b$monthcode, samples_b$sitecode, sep = "-")
# additional field to be consistent with c and d below
samples_b$sitecode_mc <- samples_b$sitecode

sum(!samples_b$sitecode %in% sites$sitecode) # 0 sitecodes consistent with sites table
sum(!samples_b$smpcode_f %in% miseq_samples$smpcode) # 0: good - all 137 samples_a$smpcode_f in miseq_samples.
sum(!samples_b$sitecode %in% sites$sitecode) # 0
sum(!samples_b$extraction_name %in% miseq_samples$extraction_name) # 0 (so sites, samples_a, and miseq_samples, consistent)

c <- as.data.frame(readxl::read_excel("~/uomShare/wergData/DNA_Barcoding/DNA barcoding sample processing Spr2022.xlsx"))
samples_c <- data.frame(sitecode  = c$sitecode, 
                        smpcode_f = NA, # No sample code field in c
                        date = c$`Date sampled`,
                        monthcode = NA,
                        extraction_name = c$`Extraction name`,
                        n_riffle = c$`number of riffle`,
                        comment = "collected by G Hehir, Unimelb",
                        reach = c$reach)
# 10 is.na n_riffles, but all included riffles according to 'sampling comments'
samples_c$n_riffle[is.na(samples_c$n_riffle)] <- 1
# remove 6 empty rows
samples_c <- samples_c[!is.na(samples_c$reach),]
samples_c$monthcode <- calcMonthDate(samples_c$date)
# 14 records with no extraction name: not collected, therefore remove
samples_c <- samples_c[!is.na(samples_c$extraction_name),]
sum(!samples_c$extraction_name %in% miseq_samples$extraction_name)  # 1 inconsistent - a typo
samples_c$extraction_name[samples_c$extraction_name == "S22KRY738"] <- "S22KRY758"
# so, as there is no sample code field in c, make smpcode_f equal to smpcodes in miseq_samples
samples_c$smpcode_f <- miseq_samples$smpcode_mc[match(samples_c$extraction_name, miseq_samples$extraction_name)]
samples_c$smpcode <- samples_c$smpcode_f
# These are incorrect, using only reach codes, and these are based on mwstr v1.2 sitecodes...
samples_c$reach <- substr(samples_c$sitecode, 1, nchar(samples_c$sitecode)-1)
# all but five of these are in the mwbugs database, and the other 5 are just minor differences
diff_reaches <-  c("CVD_644" = "CVD_663", "EDG_714" = "EDG_720", "LON_611" = "LON_598",
                   "M3D_215" = "M3D_191","TN2_118" = "TN2_104")
samples_c$reach[match(names(diff_reaches), samples_c$reach)] <- diff_reaches
sum(!samples_c$reach %in% sites$reach_v12) # 0
# so, correct sitecodes
samples_c$sitecode_mc <- samples_c$sitecode
samples_c$sitecode <- sites$sitecode[match(samples_c$reach, sites$reach_v12)]
samples_c$smpcode <- paste(samples_c$monthcode, samples_c$sitecode, sep = "-")

sum(!samples_c$sitecode %in% sites$sitecode) # 2 sitecodes consistent with sites table
samples_c$sitecode[match(c("396-MUC-987","395-TNK-1992"), samples_c$smpcode_f)] <- c("MUC-987-0","TNK-2042-3")
samples_c$smpcode[match(c("396-MUC-987","395-TNK-1992"), samples_c$smpcode_f)] <- c("396-MUC-987-0","396-TNK-2042-3")
samples_c$reach[match(c("396-MUC-987","395-TNK-1992"), samples_c$smpcode_f)] <- c("MUC-987","TNK-2042")
sum(!samples_c$smpcode_f %in% miseq_samples$smpcode) # 0: good - all 137 samples_a$smpcode_f in miseq_samples.
sum(!samples_c$extraction_name %in% miseq_samples$extraction_name) # 0 (so sites, samples_a, and miseq_samples, consistent)

d <- as.data.frame(readxl::read_excel("~/uomShare/wergData/DNA_Barcoding/DNA barcoding sample processing Aut2023.xlsx"))
# Just the ten repeated samples - note FER_648f = FER_648c (more or less!)
samples_d <- data.frame(sitecode  = d$sitecode, 
                        smpcode_f = NA, # No sample code field in c
                        date = d$`Date sampled`,
                        monthcode = NA,
                        extraction_name = d$`Extraction name`,
                        comment = "collected by G Hehir, Unimelb",
                        n_riffle = d$`number of riffle`,
                        reach = d$reach)
# 2 without dates collected by GHD - discerned the date from their supplied file 
# "~/uomShare/wergProj/bugDatabase/data_as_supplied/2022-23 MW Macroinvertebrate Data - Board Reporting.xlsx"
samples_d$date[samples_d$sitecode %in% c("YAR_3589h", "FEH_4184d")] <- as.Date("2023-05-10")
samples_d$monthcode <- calcMonthDate(samples_d$date)
samples_d$comment[d$`Sample by` == "GHD"] <- "collected by GHD"
sum(!samples_d$extraction_name %in% miseq_samples$extraction_name)  # 10 all inconsistent!
unique(miseq_samples$extraction_name[grep("A23",miseq_samples$extraction_name)])
# Not even close to consistent! Just use a manual switch
incons_ens <- c("A23BNY29042" = "A23BNY2956",
                "A23DNG318813" = "A23DNG13881",
                "A23DNG79574" = "A23DNG7957",
                "A23FEH41843" = "A23FEH4184",
                "A23FER6475" =  "A23FER648",
                "A23GRD75620" = "A23GRD7570",
                "A23TOO43342" = "A23TOO4269",
                "A23UYT1993" =  "A23UYT199",
                "A23WMI232" = "A23WM1244",
                "A23YAR36377" = "A23YAR3589")
samples_d$extraction_name[match(incons_ens, samples_d$extraction_name)] <- names(incons_ens)
sum(!samples_d$extraction_name %in% miseq_samples$extraction_name)  # 0: consistent now
# so, as there is no sample code field in d, make smpcode_f equal to smpcodes in miseq_samples
samples_d$smpcode_f <- miseq_samples$smpcode_mc[match(samples_d$extraction_name, miseq_samples$extraction_name)]
samples_d$smpcode <- samples_d$smpcode_f
# These are incorrect, using only reach codes, and these are based on mwstr v1.2 sitecodes...
samples_d$reach <- substr(samples_d$sitecode, 1, nchar(samples_d$sitecode)-1)
sum(!samples_d$reach %in% sites$reach_v12) # 0: all in the database
# so, correct sitecodes
samples_d$sitecode_mc <- samples_d$sitecode
samples_d$sitecode <- sites$sitecode[match(samples_d$reach, sites$reach_v12)]
samples_d$smpcode <- paste(samples_d$monthcode, samples_d$sitecode, sep = "-")
# 2 is.na n_riffle - not clear from table d, but these sites included riffles in c, b and a
samples_c$n_riffle[samples_c$sitecode %in% samples_d$sitecode[is.na(samples_d$n_riffle)]]
samples_d$n_riffle[is.na(samples_d$n_riffle)] <- 1

sum(!samples_d$sitecode %in% sites$sitecode) # 0 sitecodes consistent with sites table
sum(!samples_d$smpcode_f %in% miseq_samples$smpcode) # 0: good - all samples_a$smpcode_f in miseq_samples.
sum(!samples_d$extraction_name %in% miseq_samples$extraction_name) # 0 (so sites, samples_a, and miseq_samples, consistent)

samples_miseq <- rbind(samples_a, samples_b[,match(names(samples_a),names(samples_b))],
                     samples_c[,match(names(samples_a),names(samples_c))],
                     samples_d[,match(names(samples_a),names(samples_d))])

# Make samples table ready for import into mwbugs database
new_samples <- as.data.frame(matrix(nrow = nrow(samples_miseq), ncol = ncol(samples)))
names(new_samples) <- names(samples)
for(i in 1:ncol(samples_miseq)){
  if(names(samples_miseq)[i] %in% names(samples)) 
    new_samples[names(samples) == names(samples_miseq)[i]] <- samples_miseq[,i]
}
new_samples$sitecode_v12 <- sites$sitecode_v12[match(new_samples$sitecode, sites$sitecode)]
new_samples$sourcecode <- 66
new_samples$processing_method <- processing_methods$processing_method[processing_methods$pcode == "P"] 
  # i.e. "two samples combined, homogenized, DNA-metabarcoded"
new_samples$collection_method <- "RBA edge-riffle combined"
new_samples$collection_method[samples_miseq$n_riffle == 0] <- "RBA two edges combined"
new_samples$comment = "count value is number of records from 3 miseq replicates"
new_samples$comment[new_samples$smpcode == "384-ANS-1611-6-DP"] <- "count value is number of records from 2 miseq replicates"
new_samples$embargoed <- 1
new_samples$old_samplecode <- samples_miseq$smpcode_f
new_samples$dry_no_sample <- 0  
# [CHRIS go back and include samples not taken because of dryngess - check 'Number of edge' field for hints]
smpcode_vec <- new_samples$smpcode
new_samples$smpcode <- paste0(smpcode_vec, "-DP")
new_samples$smpcode[samples_miseq$n_riffle == 0] <- paste0(smpcode_vec[samples_miseq$n_riffle == 0], "-EP")

# Add the new sourcecode to the mwbugs spv table.
new_spv <- data.frame(sourcecode = 66, project = "ARC LP DNA metabarcoding", reference = NA,
                      e_source = "github.com/mecarew/metabarcoding_workflow/src_water_prod/Appendix3_import_miseq_to_mwbugs.qmd",
                      file_locations = 
"wergStaff/MelCarew/git-data/metabarcoding_workflow",
lab = "PEARG and WERG Unimelb, led by Melissa Carew",
comment = NA)
```

Correct taxonomy names prior to matching to the database

```{r}

# Correct spelling errors
bugdata_spp$species[bugdata_spp$species == "Austrochiltonia sp. MC1"] <- "Austrochiltonia sp. MC-1"
bugdata_spp$species[bugdata_spp$species == "Tamasia acuta/Tamasia variegata"] <- "Tamasia acuta/variegata"
bugdata_spp$species <- gsub("ParaLimnophila","Paralimnophila",bugdata_spp$species)
bugdata_spp$genus[bugdata_spp$genus == "ParaLimnophila"] <- "Paralimnophila"
bugdata_spp$species <- gsub("Aeschnidae","Aeshnidae",bugdata_spp$species)
bugdata_spp$species[bugdata_spp$species == "Nososticta solitaris"] <- "Nososticta solitaria"
bugdata_spp$family[bugdata_spp$family == "Trapezidae"] <- "Trapeziidae"
# Some Staphylinidae with order = Diptera
bugdata_spp$order[bugdata_spp$family == "Staphylinidae"] <- "Coleoptera"
#[MEL: Please check the following three sets of decisions to exclude.  Do you agree?]
# 5 family-level records identified as Unionidae. Genbank match gastropod ~92% - delete
bugdata_fam <- bugdata_fam[bugdata_fam$family != "Unionidae",]
# 6 family-level records identified as Unionidae.  Genbank match gastropod ~92% - delete
bugdata_fam <- bugdata_fam[bugdata_fam$family != "Arcidae",]
# Unidentified higher-order records of unknown (likely terrestrial) ecology.  Exclude  
bugdata_spp <- bugdata_spp[!bugdata_spp$species %in% c("Oligochaeta sp. MC-3",
          "Amphipoda sp. B-AFN5257","Isopoda sp. B-ADV3337",
          "Chrysomelidae sp. B-AAQ0004","Lumbricidae sp. MC-1",
          "Galerucinae sp. B-AET7210","Ephydridae sp. MC-3",
          "Aturidae sp. B-AEH7912","Dolichopodidae sp. B-ADP0042",
          "Muscidae sp. B-AAU5065","Syllidae sp. MC-O1",
          "Melitidae sp. MC-1","Galerucinae sp. B-ABA9962",
          "Coleoptera sp. B-AEI6253","Muscidae sp. B-AES9495/AFJ0397",
          "Ephydridae sp. B-ACZ3069","Galerucinae sp. B-ABX9821",
          "Osmylidae sp. MC-1","Stratiomyidae sp. MC-SC1",
          "Pionidae sp. B-ACW4137"),]

# Check all families are in the mwbugs database
all_fams <- unique(c(bugdata_spp$family, bugdata_gen$family, bugdata_fam$family))
non_match_fams <- data.frame(family = all_fams[!(all_fams %in% taxon_fam$family | all_fams %in% taxon_fam$subfamily)], 
                             order_mc = NA, order_gbif = NA)
non_match_fams <- non_match_fams[non_match_fams$family != "",]  # zero non-matching!

# Check all genera are in the mwbugs database
all_genera <- unique(c(bugdata_spp$genus, bugdata_gen$genus, bugdata_fam$genus))
non_match_gen <-  all_genera[!all_genera %in% taxon_gen$genus]
non_match_gen <- non_match_gen[non_match_gen != ""]  # zero non-matching!
```


The following identifies new species to be added to the database.  As the 46-site data was added prior to finalization of species groups, I propose removing the 46-site bugdata altogether, and removing any "B-" species names that ended up being folded into species groups...

I need a final synonym table for that... not yet complete. But I will prepare the code now for adding new species to the database in preparation for that finalisation.

```{r}

all_spp <- unique(bugdata_spp$species)
new_spp <- all_spp[!all_spp %in% c(taxon_spp$species, morphospp_etc$lowesttaxon)]

# CW manually checked all form_spp names against mwbugs taxon_spp. 
form_spp <- new_spp[!grepl("MC-",new_spp) & !grepl("B-",new_spp) & !grepl(" group", new_spp)]
form_spp <- form_spp[order(form_spp)]
# All were new except:
# Synonym to add to mwbugs
# QO120201 Aeshna brevistyla = Adversaeschna brevistyla
# Make "Austroargiolestes spp." QD070200

new_spp <- data.frame(bugcode = NA, species = new_spp, mwbugs_table = NA)
for(i in 1:nrow(new_spp)){
  spi <- new_spp$species[i]
  geni <- strsplit(spi, " ")[[1]][1]
  gencode_i <- taxon_gen$gencode[taxon_gen$genus == geni]
  all_geni_codes <- c(taxon_spp$spcode[taxon_spp$gencode == gencode_i],
                      morphospp_etc$taxoncode[substr(morphospp_etc$lowesttaxon,1,nchar(geni)) == geni])
  used_suffixes <- substr(all_geni_codes,7,8)
  if(geni %in% taxon_gen$genus){
  if(!grepl("MC-",spi) & !grepl("B-",spi) & !grepl(" group", spi) & !grepl("/", spi)){
 # If it is a formal species name, give it the next numeric code.
     used_suffixes <- used_suffixes[used_suffixes %in% all_prs$prs[all_prs$number == 1]]
     new_spp$bugcode[i] <- paste0(gencode_i, all_prs$prs[all_prs$number == 1 & 
                                                     !all_prs$prs %in% used_suffixes][1])
     new_spp$mwbugs_table[i] <- "taxon_spp"
  }else{
  # or else, give it the next alphanumeric code.
     used_suffixes <- used_suffixes[used_suffixes %in% all_prs$prs[all_prs$number == 0]]
     new_spp$bugcode[i] <- paste0(gencode_i, all_prs$prs[all_prs$number == 0 & 
                                                     !all_prs$prs %in% used_suffixes][1])
    # if the species is identified by a single BOLD identifier, then it can go in taxon_spp
     new_spp$mwbugs_table[i] <- ifelse(!grepl("MC-",spi) & !grepl(" group", spi) & !grepl("/", spi),
                                       "taxon_spp", "morphospp_etc")
  }
  # 
 if(spi == "Austroargiolestes spp.") {
   new_spp$bugcode[i] <- "QD070200"
   new_spp$mwbugs_table[i] <- NA
 }
  }else{
      #If species only identified to family or subfamily...
    # First take care of some non-listed taxon assignments
    non_listed <- c("Atalophlebiinae" = "Leptophlebiidae",
                    "Limnophilinae" = "Limoniidae")
    if(geni %in% names(non_listed)){
      geni <- unname(non_listed[names(non_listed) == geni])
    }
    if(geni %in% taxon_fam$family){
      famcodei <- taxon_fam$famcode[taxon_fam$family == geni]
      if(geni == "Chironomidae") famcodei <- "QDAZ"
    }else{
    if(geni %in% taxon_fam$subfamily){
      famcodei <- taxon_fam$famcode[!is.na(taxon_fam$subfamily) & taxon_fam$subfamily == geni]
      if(geni == "Chironominae") famcodei <- "QDAJ"
    }
    }
    allfam99i_codes <- taxon_all$shortcode[substr(taxon_all$shortcode,1,6) == 
                                             paste0(famcodei,"99")]
    used_suffixes <- substr(allfam99i_codes,7,8)
    used_suffixes <- used_suffixes[used_suffixes %in% all_prs$prs[all_prs$number == 0]]
    new_spp$bugcode[i] <- paste0(famcodei, "99", all_prs$prs[all_prs$number == 0 & 
                                                     !all_prs$prs %in% used_suffixes][1])
    new_spp$mwbugs_table[i] <- "morphospp_etc"
    }
}

# Subsequent checks found one 'species', Austropsyche sp., should actually be 
# Austropsyche  sp. B-AAW3080, and its asv_code was incorrectly called and unidentified 
# "Psychidae in miseq15.  This asv_code needs to be added to the next version of the 
# asv_library, but for now, we will apply a manual fix to the workflow results here.
bugdata_spp <- rbind(bugdata_spp, bugdata_fam[bugdata_fam$asv_code == 
                                                "00f4bc36686affa317cbbdbad7a65991",])
bugdata_fam <- bugdata_fam[bugdata_fam$asv_code != 
                                                "00f4bc36686affa317cbbdbad7a65991",]
bugdata_spp$genus[bugdata_spp$asv_code == "00f4bc36686affa317cbbdbad7a65991"] <- "Austropsyche"
bugdata_spp$species[bugdata_spp$asv_code == "00f4bc36686affa317cbbdbad7a65991"] <- 
  "Austropsyche sp. B-AAW3080"
new_spp$species[new_spp$species == "Austropsyche sp."] <- "Austropsyche sp. B-AAW3080"
new_spp$bugcode[new_spp$species == "Austropsyche sp. B-AAW3080"] <- "QT0603AI"

# Note that morphospp_etc contains QT0603AI Austropsyche sp. B-AAW3080 (from 46-site data)
# I will move all such records from morphospp_etc to taxon_spp to be consistent with 
# the convention used here.

new_taxon_spp <- data.frame(gencode = substr(new_spp$bugcode,1,6),
                            spcode = new_spp$bugcode,
                            species = new_spp$species,
                            type = "DNA barcoded taxon with multiple bins or without confirmed ID",
                            tempNote = "added by CW for ARC metabarcoding project",
                            t1code = "")[!is.na(new_spp$mwbugs_table) &
                                           new_spp$mwbugs_table == "taxon_spp",]

new_morphospp_etc <- data.frame(taxon_code = new_spp$bugcode,
                                lowesttaxon= new_spp$species,
                                tempNote = "added by CW for ARC metabarcoding project",
                                taxonomic_note = substr(new_spp$bugcode,1,2))[!is.na(new_spp$mwbugs_table) &
                                                                                new_spp$mwbugs_table == "morphospp_etc",]
```

Assemble biological data for import into database

```{r}
# # some max_p_identity inconsistencies
# x <- unique(bugdata_spp[c("asv_code","max_p_identity","amplicon")])
# x <- x[duplicated(x$asv_code) | duplicated(x$asv_code, fromLast = TRUE),]
# x[order(x$asv_code),]

# Ad hoc corrections to 3 inconsistent max_p_identity values (the first concerning
# "Austropsyche sp. B-AAW3080", above)
bugdata_spp$max_p_identity[bugdata_spp$asv_code == "00f4bc36686affa317cbbdbad7a65991"] <- 99.7
bugdata_spp$max_p_identity[bugdata_spp$asv_code == "bfb774b22881eee84f0f221daeb9d157"] <- 100
bugdata_spp$max_p_identity[bugdata_spp$asv_code == "ed6b0b7d5ab0813489018124e4f46c1e"] <- 99.7

bugdata_spp$extraction_name <- substr(bugdata_spp$sample,1,nchar(bugdata_spp$sample)-4)
bugdata_spp$smpcode <- samples_miseq$smpcode[match(bugdata_spp$extraction_name, 
                                                   samples_miseq$extraction_name)]
# The NAs introduced here are the three samples (1418 records) with smpcodes starting with 337 excluded above
sequences <- unique(bugdata_spp[c("asv_code","asv_seq","max_p_identity","amplicon")])

biota_wrkng <- bugdata_spp[c("sample","smpcode","asv_code","species")]
biota_wrkng <- biota_wrkng[!is.na(biota_wrkng$smpcode),]
biota_wrkng$bugcode <- taxon_all$shortcode[match(biota_wrkng$species, taxon_all$taxon)]
for(i in which(is.na(biota_wrkng$bugcode))){
    biota_wrkng$bugcode[i] <- new_spp$bugcode[new_spp$species == biota_wrkng$species[i]]
   }
sample_asv <- unique(biota_wrkng[c("asv_code","smpcode")])
bugcode_asv <- unique(biota_wrkng[c("asv_code","bugcode")])

biota_agg <- aggregate(biota_wrkng$bugcode, by = list(smpcode = biota_wrkng$smpcode,
                                                      taxoncode = biota_wrkng$bugcode,
                                                      species = biota_wrkng$species), FUN = length)

new_biota <- data.frame(smpcode = biota_agg$smpcode,
                        taxoncode = biota_agg$taxoncode,
                        count = biota_agg$x,
                        taxon = biota_agg$species, 
                        coarsepick = 0, 
                        notes = "",
                        originalbugcode = NA,
                        shortcode = NA)

# Add records only identified to genus:
bugdata_gen$extraction_name <- substr(bugdata_gen$sample,1,nchar(bugdata_gen$sample)-4)
bugdata_gen$smpcode <- samples_miseq$smpcode[match(bugdata_gen$extraction_name, 
                                                   samples_miseq$extraction_name)]
sequences <- rbind(sequences,
                   unique(bugdata_gen[c("asv_code","asv_seq","max_p_identity","amplicon")]))
# NAs introduced as above
biota_wrkng <- bugdata_gen[c("sample","smpcode","asv_code","genus")]
biota_wrkng <- biota_wrkng[!is.na(biota_wrkng$smpcode),]
biota_wrkng$bugcode <- paste0(taxon_gen$gencode[match(biota_wrkng$genus, taxon_gen$genus)],"99")

biota_agg <- aggregate(biota_wrkng$bugcode, by = list(smpcode = biota_wrkng$smpcode,
                                                      taxoncode = biota_wrkng$bugcode,
                                                      species = biota_wrkng$genus), 
                                                      FUN = length)
biota_agg$species <- paste0(biota_agg$species, " sp. (Unident.)")

new_biota <- rbind(new_biota, data.frame(smpcode = biota_agg$smpcode,
                        taxoncode = biota_agg$taxoncode,
                        count = biota_agg$x,
                        taxon = biota_agg$species, 
                        coarsepick = 0, 
                        notes = "",
                        originalbugcode = NA,
                        shortcode = NA))

sample_asv <- rbind(sample_asv, unique(biota_wrkng[c("asv_code","smpcode")]))
bugcode_asv <- rbind(bugcode_asv, unique(biota_wrkng[c("asv_code","bugcode")]))

# Add records only identified to family:
bugdata_fam$extraction_name <- substr(bugdata_fam$sample,1,nchar(bugdata_fam$sample)-4)
bugdata_fam$smpcode <- samples_miseq$smpcode[match(bugdata_fam$extraction_name, 
                                                   samples_miseq$extraction_name)]
# NAs introduced as above
sequences <- rbind(sequences,
                   unique(bugdata_fam[c("asv_code","asv_seq","max_p_identity","amplicon")]))
biota_wrkng <- bugdata_fam[c("sample","smpcode","asv_code","family")]
biota_wrkng$extraction_name <- substr(biota_wrkng$sample,1,nchar(biota_wrkng$sample)-4)
biota_wrkng$smpcode <- samples_miseq$smpcode[match(biota_wrkng$extraction_name, 
                                                   samples_miseq$extraction_name)]
biota_wrkng <- biota_wrkng[!is.na(biota_wrkng$smpcode),]
biota_wrkng$bugcode <- paste0(taxon_fam$famcode[match(biota_wrkng$family, taxon_fam$family)],"9999")
biota_agg <- aggregate(biota_wrkng$bugcode, by = list(smpcode = biota_wrkng$smpcode,
                                                      taxoncode = biota_wrkng$bugcode,
                                                      species = biota_wrkng$family), 
                                                      FUN = length)
biota_agg$species <- paste0(biota_agg$species, " sp. (Unident.)")
new_biota <- rbind(new_biota, data.frame(smpcode = biota_agg$smpcode,
                        taxoncode = biota_agg$taxoncode,
                        count = 1,
                        taxon = biota_agg$species, 
                        coarsepick = 0, 
                        notes = "",
                        originalbugcode = NA,
                        shortcode = NA))

sample_asv <- rbind(sample_asv, unique(biota_wrkng[c("asv_code","smpcode")]))
bugcode_asv <- rbind(bugcode_asv, unique(biota_wrkng[c("asv_code","bugcode")]))

#####
x <- bugcode_asv[duplicated(bugcode_asv$asv_code) | duplicated(bugcode_asv$asv_code, fromLast = TRUE),]
# 0 asv_codes with > 1 identification

new_biota <- new_biota[order(new_biota$smpcode),]
```


```{r}
bins <- as.data.frame(readxl::read_excel("data/asv_library_corrected.xlsx", sheet = 3))
bins <- bins[!is.na(bins$bin_uri),]

miseq_data <- rbind(bugdata_spp[c("sample","asv_code","reads")],
                    bugdata_gen[c("sample","asv_code","reads")],
                    bugdata_fam[c("sample","asv_code","reads")])
miseq_data <- dplyr::mutate(miseq_data, 
                            smpcode = miseq_data$smpcode[match(miseq_data$sample, 
                                                               miseq_samples$seq_code)],
                            .after = sample)

 reads_tab <- dplyr::mutate(miseq_data, 
                            smpcode = miseq_samples$smpcode[match(miseq_data$sample, miseq_samples$seq_code)],
                            bugcode = bugcode_asv$bugcode[match(miseq_data$asv_code, bugcode_asv$asv_code)], 
                            .after = sample)
 reads_tab <- dplyr::mutate(reads_tab, 
                            taxon = new_biota$taxon[match(paste0(reads_tab$smpcode,reads_tab$bugcode),
                                                          paste0(new_biota$smpcode,new_biota$taxoncode))],
                            .after = bugcode)
```

```{r}
#Build xlsx file for inspection prior to adding these data to the mwbugs database
table_list <- list(sites = sites[sites$sitecode %in% unique(new_samples$sitecode),],
                   new_samples = new_samples,
                   new_biota = new_biota,
                   collection_methods = collection_methods[collection_methods$ccode %in% c("D","E"),],
                   processing_methods = processing_methods[processing_methods$pcode %in% c("P"),],
                   new_spv = new_spv,
                   sequences = sequences,
                   bugcode_asv = bugcode_asv, 
                   sample_asv = sample_asv,
                   new_taxon_spp = new_taxon_spp,
                   new_morphospp_etc = new_morphospp_etc, 
                   miseq_samples = miseq_samples,
                   miseq_data = miseq_data,
                   bins = bins)
WriteXLS::WriteXLS(table_list, "~/temp/arc_metabarcoding_data_for_mwbugs_db.xlsx")
```
